---
title: "Digitec Analysis"
format:
  html:
    df-print: paged
    toc: true
    code-tools: true

editor: visual
author: Fereshteh Ahmadi, Anna Barcikowska, Vanmassenhove Sen Candice, Colin Marti
---

## 1. Business Plan Analysis

Digitec Galaxus is Switzerland’s largest online retailer for consumer electronics, operating exclusively online. While the company currently runs nationwide campaigns, our analysis shows that customer behaviour varies significantly across regions and demographics. These differences matter most for high-impact categories like notebooks, which drive a large share of revenue, and for strategic niche segments such as gaming products, which, although smaller in size, offer strong engagement and brand-building opportunities.

This project explores how these demographic and regional factors influence online purchases using: - \~163,000 Digitec transaction records, and - Canton-level demographic data from the Swiss Federal Statistical Office (BFS).

The findings will guide a data-driven marketing strategy to personalize digital campaigns and increase online sales efficiency.

### 1.1. Business Problem

> **How can Digitec leverage demographic and regional differences to optimise marketing strategies for high-impact categories, while also tailoring campaigns for strategic niche categories?**

Digitec handles thousands of sales every day across many product categories, regions and customer types. But most of the marketing is still done on a national level – one broad message for the whole country. That raises a simple question: Are all customers in Switzerland really the same, or do different regions buy differently? If purchasing behaviour varies by canton or by demographics, then Digitec might be spending budget on campaigns that are not equally effective everywhere. That's why we decided to dive into this problem and explore high- and low-impact categories of Digitec products.

### 1.2. Data Overview

| Source | Description | Key Variables |
|:-----------------------|:-----------------------|:-----------------------|
| **Digitec Transactions (`DigitecLive_Final.csv`)** | \~163k web-scraped online transactions from Digitec’s live purchase feed | `brandName`, `infos.Category`, `salesPrice.amountIncl`, `dateTime`, `mun_canton` |
| **BFS Population Data (`population_2023.xlsx`)** | Canton-level demographics (2023) | `Total`, `Swiss`, `Foreigner`, `Urban`, `Intermediate`, `Rural`, `Age_0_19`, `Age_20_64`, `Age_65plus`, `Single`, `Married`, `Widowed`, `Divorced` |
| **BFS Population by Age (`mns-population-by-age-2023.xlsx`)** | Detailed population structure by age and region | `Region`, `Total`, `Age_0_4`, `Age_5_9`, …, `Age_80plus` |
| **BFS Population by Gender (`mns-population-by-gender-2023.xlsx`)** | Gender composition by canton and municipality (2023) | `Region`, `Total_All`, `Mann_All`, `Frau_All`, `Swiss`, `Foreigner` |
| **Swiss Municipal Boundaries (`swissBOUNDARIES3D_1_5_TLM_HOHEITSGEBIET.shp`)** | Official Swiss municipality polygons used to map Digitec transactions geographically and join by BFS municipality number | `BFS_NUMMER` (municipality ID), `NAME` (municipality name), `KANTONSNUM` (canton ID), `geometry` (MULTIPOLYGON) |

These additional datasets allow the inclusion of **more granular demographic information** - enabling a refined analysis of age and gender-related purchasing behavior across regions.

### 1.3. Unit of Analysis

The analysis is now conducted at the **municipality level** ($n \approx 2,200$), using the official **BFS municipality number (`mun_bfsnr`)** as the join key between datasets.

Each municipality represents one observation containing: - $sales\_per\_capita = \frac{total\_sales}{population}$\
- Average order value (CHF)\
- Age-group shares (e.g., % under 20, % over 65)\
- Gender composition (% male, % female)\
- Citizenship composition (% Swiss, % foreign)

This finer-grained approach allows for **more localized insights**, capturing demographic and purchasing differences that are hidden when aggregating by canton.

------------------------------------------------------------------------

## 2. Setup and Data Loading

This first code chunk sets up the R environment by loading all necessary libraries for the analysis and loading the four primary datasets. You can have a look only if you want.

```{r}
#| label: setup
#| code-fold: true
#| message: false
#| warning: false

# Load all required libraries for the project
library(tidyverse) 
library(readxl)    
library(lubridate) 
library(stringr)   
library(dlookr)
library(ggrepel)
library(ggcorrplot)
library(ggplot2)
library(viridis)
library(viridisLite)
library(dplyr)
library(plotly)
library(sf)
library(scales)
library(tidyr)
library(lubridate)
library(forcats)




# --- GLOBAL DESIGN DEFINITION ---
# Save the original theme_minimal
orig_theme_minimal <- ggplot2::theme_minimal

# Override with custom settings
theme_minimal <- function(base_size = 11, base_family = "", ...) {
  orig_theme_minimal(base_size = base_size, base_family = base_family, ...) +
  theme(
    # 1. Axis text (values on ticks)
    axis.text = element_text(size = 18, color = "black"),
    
    # 2. Axis titles 
    axis.title = element_text(size = 20, color = "black", margin = margin(t=15, r=15)),
    
    # 3. Main Plot Title
    plot.title = element_text(size = 26, color = "black", margin = margin(b=15)),
    
    # 4. Plot Subtitle
    plot.subtitle = element_text(
      size = 20, 
      color = "black", 
      margin = margin(b=20)
    ),
    
    # 5. Caption 
    plot.caption = element_text(size = 14, color = "grey30", hjust = 1),
    
    # 6. Legends
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18)
  )
}

# Load digitect dataset
load("digitec.RData") # Loads the 'data' object

# Save to descriptive variable
digitec_data <- get("data")

## Load BFS data sets
canton_population <- read_excel("population_2023.xlsx")
municipal_population_gender <- read_excel("mns-population-by-gender-2023.xlsx")
municipal_population_age <- read_excel("mns-population-by-age-2023.xlsx")

```

## 3. Data Cleaning and Preprocessing

This section details the systematic process of cleaning and preparing the raw data of the required data sets. The goal is to produce a reliable, aggregated dataset at the municipal and cantonal level. The municipal data will be a combination of three datasets: "mns-population-by-age", "mns-population-by-gender", and the digitec dataset. The cantonal dataset will be a merging of the BFS demographic data and the digitec dataset.

### 3.1. Digitec Dataset

An initial inspection of the raw Digitec data revealed several necessary cleaning steps:

-   **Correcting Data Types:** The `dateTime` column was stored as a character string and needed to be converted into a proper date-time format for time-based analysis.
-   **Removing Irrelevant Columns:** Many columns were redundant (e.g., price with and without VAT), contained technical metadata, or were not relevant to the analysis (e.g., image URLs).
-   **Ensuring Data Quality:** Transactions without a price or a clear, unique geographic match (`mun_nhits != 1`) were removed to ensure the reliability of the analysis.
-   **Feature Engineering:** More useful features were extracted from the raw data, such as the day of the week from the `dateTime` column or the clean product category name from a URL string.

#### 3.1.1. Data Cleaning & Engineering

First we check the dataset columns that we think might require some filtering or deletion.

```{r}

candidates <- c(
  "cityName",
  "salesPrice.currency",
  "infos.Price",
  "infos.Currency",
  "salesPrice",
  "mun_nhits"
)

existing <- intersect(candidates, names(digitec_data))

check_var <- function(df, var) {
  x <- df[[var]]
  tibble(
    variable      = var,
    exists        = TRUE,
    n             = length(x),
    n_non_na      = sum(!is.na(x)),
    prop_na       = mean(is.na(x)),
    n_unique_non_na = dplyr::n_distinct(x, na.rm = TRUE),
    example_values  = paste(utils::head(unique(x), 3), collapse = " | ")
  )
}

profile <- map_dfr(existing, ~check_var(digitec_data, .x))
profile
```

-   `salesPrice.currency`, `infos.Currency` → `n_unique_non_na == 1` (“CHF”) we decided to remove this column.
-   `salesPrice` → `n_non_na == 0` empty column so we decided to delete it.
-   `mun_nhits` → mostly/all 1

```{r}
if ("mun_nhits" %in% names(digitec_data)) {
  cat("mun_nhits frequency (percent):\n")
  perc <- prop.table(table(digitec_data$mun_nhits, useNA = "ifany")) * 100
  print(round(perc, 2))
}
```

-   **`1`** → 153 778 or 94.35% of the rows matched exactly one municipality.
-   **`0`** → 9 194 or 5.64% of the rows had *no municipality match*.
-   **`2`** → 21 or 0.01% of the rows matched *two possible municipalities*.

We would drop the rows where `mun_nhits == 0` or `mun_nhits == 2`, then remove the `mun_nhits` column. Since we already have 95%.

-   `cityName` → unique city name inserted by customer, since we have the city and municipality takes from SWISS POST Api we decided to drop this column.
-   `infos.Price` → many NAs - 63% of the data is missing so the column is invalid.

```{r}
if (all(c("infos.Price", "salesPrice.amountIncl") %in% names(digitec_data))) {
  same_info <- digitec_data %>%
    filter(!is.na(infos.Price) & !is.na(salesPrice.amountIncl)) %>%
    summarise(
      n = n(),
      max_abs_diff = max(abs(infos.Price - salesPrice.amountIncl), na.rm = TRUE)
    )
  same_info
}
```

60 577 rows had *both* `infos.Price` and `salesPrice.amountIncl` . The value `max_abs_diff = 1021.9` means that in some rows, the difference between the two price fields is CHF 1 021.90 so prices differ a lot. We decided that this column is inconsistent.

The following steps were implemented in a single, coherent `dplyr` pipeline for efficiency and reproducibility. The result is a cleaned DataFrame named `data_clean`.

```{r}
#| label: data-cleaning-pipeline

digitec_data_clean <- digitec_data %>%
  # 1) Select and rename relevant columns for better readability
  rename(
    price        = salesPrice.amountIncl,
    category_url = infos.Category,
    sales_rank_raw = infos.Sales_Rank,
    canton       = mun_canton,
    bfsnr        = mun_bfsnr
  ) %>%
  select(
    price, dateTime, brandName, fullProductName,
    socialShoppingTransactionTypeId,
    category_url, sales_rank_raw, canton, mun_nhits, bfsnr
  ) %>%
  
  # 2) Filter rows to ensure data quality
  filter(
    !is.na(price),                  # keep only transactions with a price
    !is.na(canton),                 # keep only transactions with canton info
    mun_nhits == 1,                 # keep only unique geographic matches
    socialShoppingTransactionTypeId %in% c(3, 5),  # only bought/shipped
    grepl("von", sales_rank_raw)    # keep only rows where rank contains "von"
  ) %>%
  
  # 3) Remove irrelevant or redundant columns
  select(-any_of(c(
    "cityName",
    "salesPrice.currency",
    "infos.Price",
    "infos.Currency",
    "salesPrice"
  ))) %>%
  
  # 4) Create new, useful columns (feature engineering)
  mutate(
    # timestamps
    date     = as_date(dateTime),
    dateTime = ymd_hms(dateTime),
    hour     = hour(dateTime),
    weekday  = wday(dateTime, label = TRUE, week_start = 1),
    # product category from URL
    category = basename(category_url)
  ) %>%
  
  # 5) Parse sales rank: "12 von 345" -> numeric position / total
  separate(
    sales_rank_raw,
    into = c("rank_position", "rank_total"),
    sep  = " von "
  ) %>%
  mutate(
    rank_position = as.numeric(rank_position),
    rank_total    = as.numeric(rank_total),
    rank_relative = rank_position / rank_total,
    rank_score    = 1 - rank_relative
  ) %>%
  
  # 6) Recode transaction type and convert key vars to factors
  mutate(
    transaction_type = case_when(
      socialShoppingTransactionTypeId == 3 ~ "bought",
      socialShoppingTransactionTypeId == 5 ~ "shipped"
    ),
    transaction_type = fct_relevel(transaction_type, "bought", "shipped"),
    category  = as.factor(category),
    brandName = as.factor(brandName)
  ) %>%
  select(-socialShoppingTransactionTypeId, -mun_nhits) %>%  # no longer needed
  
  # 7) Drop duplicates
  distinct() %>%
  
  # 8) Handle Inf and winsorize rank metrics in one go
  mutate(
    rank_score    = ifelse(is.infinite(rank_score),    NA_real_, rank_score),
    rank_relative = ifelse(is.infinite(rank_relative), NA_real_, rank_relative)
  ) %>%
  { 
    qs_score <- quantile(.$rank_score,    probs = c(0.025, 0.975), na.rm = TRUE)
    qs_rel   <- quantile(.$rank_relative, probs = c(0.025, 0.975), na.rm = TRUE)
    
    mutate(.,
      rank_score    = pmin(pmax(rank_score,    qs_score[1]), qs_score[2]),
      rank_relative = pmin(pmax(rank_relative, qs_rel[1]),   qs_rel[2])
    )
  } %>%
  filter(!is.na(rank_score), !is.na(rank_relative))
```

```{r}
overview(digitec_data_clean)
summary(digitec_data_clean)

```

### 3.2. BFS Datasets

The data within the three BFS is mostly clean. There are no apparent missing values or errors, just some unnecessary rows. More importantly, the datasets need to be compatible for the upcoming merge with the Digitec data.\
\
The issues that we identified were as follows:

-   **Column names:** The column names of all three datasets need to be non-capitalized and follow the snake_case format in order to align with the digitec dataset.

-   **Unwanted rows:** Remove the unwanted canton and bezirk rows with the municipality datasets and rename the column to muncipality.

-   **Canton codes:** The canton names within the municipality BFS dataset are not in the standard code format.

-   **Municipality number:** The age and gender datasets have an improper municipality notation. Later on, we'll have to merge these datasets with the digitec dataset according to the municipality number, so it's imperative that we make sure they're in the same format.

-   **Ages:** Currently ages within the age datasetare not grouped in the same way as BFS' canton dataset.

#### 3.2.1 Data Cleaning & Engineering

**Municipality Level**

Firstly, we will tackle the column names; we do this simply by running each dataset through this snake_case function:

```{r}

#function which converts input to snake_case
clean_names <- function(x) {
  x %>%
    str_to_lower() %>%
    
    # Replace spaces with underscores
    str_replace_all(" ", "_")             
}

# Rename each dataset
colnames(canton_population) <- clean_names(colnames(canton_population))
colnames(municipal_population_age) <- clean_names(colnames(municipal_population_age))
colnames(municipal_population_gender) <- clean_names(colnames(municipal_population_gender))
```

Then within the age and gender datasets we can remove the unnecessary rows, rename the column and extract the bfsnr from the municipalities.

```{r}
#| label: row-removal

# Rename first column to bfsnr like in the digitec dataset
colnames(municipal_population_age)[1] <- "bfsnr"
colnames(municipal_population_gender)[1] <- "bfsnr"

# Clean municipality age data
municipal_population_age_clean <- municipal_population_age %>%
  filter(str_detect(bfsnr, "^\\.+")) %>%  # Match rows starting with one or more dots
  mutate(
    bfsnr = str_extract(bfsnr, "\\d+"),  # Extract numeric code
    bfsnr = as.integer(bfsnr)
  )

# Clean municipality gender data
municipal_population_gender_clean <- municipal_population_gender %>%
  filter(str_detect(bfsnr, "^\\.+")) %>%
  mutate(
    bfsnr = str_extract(bfsnr, "\\d+"),
    bfsnr = as.integer(bfsnr)
  )

municipal_population_age_clean
municipal_population_gender_clean

```

Now we group the ages from 10-19, 20-64, and 65-100+ and our age and population datasets are ready for merging.

```{r}

# Collapse into age bands
# Collapse into 10-year age bands at municipality level
municipal_population_age_clean <- municipal_population_age_clean %>%
  rowwise() %>%
  mutate(
    age_10_19  = sum(c_across(`10`:`19`), na.rm = TRUE),
    age_20_29  = sum(c_across(`20`:`29`), na.rm = TRUE),
    age_30_39  = sum(c_across(`30`:`39`), na.rm = TRUE),
    age_40_49  = sum(c_across(`40`:`49`), na.rm = TRUE),
    age_50_59  = sum(c_across(`50`:`59`), na.rm = TRUE),
    age_60_69  = sum(c_across(`60`:`69`), na.rm = TRUE),
    age_70_79  = sum(c_across(`70`:`79`), na.rm = TRUE),
    age_80_plus = sum(c_across(`80`:`99`), `100_und_mehr`, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  select(
    bfsnr, total,
    age_10_19, age_20_29, age_30_39,
    age_40_49, age_50_59, age_60_69, age_70_79, age_80_plus
  )


```

**Canton Level**

For the canton level we need to map the canton names found in the canton BFS dataset to their official abbreviations used in the transaction dataset.

```{r}
#| label: canton-code-mapping


# Define canton name-to-code mapping
canton_codes <- list(
  "Zurich" = "ZH",
  "Bern" = "BE",
  "Lucerne" = "LU",
  "Uri" = "UR",
  "Schwyz" = "SZ",
  "Obwalden" = "OW",
  "Nidwalden" = "NW",
  "Glarus" = "GL",
  "Zug" = "ZG",
  "Fribourg" = "FR",
  "Solothurn" = "SO",
  "Basel-Stadt" = "BS",
  "Basel-Landschaft" = "BL",
  "Schaffhausen" = "SH",
  "Appenzell A. Rh." = "AR",
  "Appenzell I. Rh." = "AI",
  "St. Gallen" = "SG",
  "Graubünden" = "GR",
  "Aargau" = "AG",
  "Thurgau" = "TG",
  "Ticino" = "TI",
  "Vaud" = "VD",
  "Valais" = "VS",
  "Neuchâtel" = "NE",
  "Geneva" = "GE",
  "Jura" = "JU"
)

# Convert to named vector for mapping
code_map <- setNames(unlist(canton_codes), names(canton_codes))

# Replace canton names with codes and remove non-cantons
cleaned_canton_population <- canton_population %>%
  filter(canton %in% names(canton_codes)) %>%
  mutate(canton = code_map[canton])


# Display the processed excel data
cleaned_canton_population

```

## 4. Validation and Final Aggregation

### 4.1 Digitec Dataset

An initial aggregation of the cleaned data surprisingly resulted in 29 groups instead of the 26 expected Swiss cantons. An investigation of the `canton` column revealed that the dataset also contained transactions from neighboring countries, which needed to be excluded for this analysis.

```{r}
#| label: validation-step

# A list of valid Swiss canton abbreviations
valid_swiss_cantons <- c(
  "AG", "AI", "AR", "BE", "BL", "BS", "FR", "GE", "GL", "GR", 
  "JU", "LU", "NE", "NW", "OW", "SG", "SH", "SO", "SZ", "TG", 
  "TI", "UR", "VD", "VS", "ZG", "ZH"
)

# Identify which unique codes in our data are NOT in the valid list
non_swiss_codes <- setdiff(unique(digitec_data_clean$canton), valid_swiss_cantons)

print("Found the following non-Swiss location codes:")
non_swiss_codes
```

Based on this finding, the dataset was filtered to include only transactions from the 26 Swiss cantons. Afterward, the final aggregation was performed to create the analysis-ready dataset.

```{r}
#| label: final-aggregation

# Filter the dataset to retain only transactions from Swiss cantons
data_final_swiss <- digitec_data_clean %>%
  filter(canton %in% valid_swiss_cantons)

# Perform the final aggregation at the cantonal level
canton_data_final <- data_final_swiss %>%
  group_by(canton) %>%
  summarise(
    total_sales = sum(price, na.rm = TRUE),
    num_transactions = n(),
    avg_order_value = mean(price, na.rm = TRUE),
    top_brand = names(which.max(table(brandName))),
    top_category = names(which.max(table(category)))
  ) %>%
  arrange(desc(total_sales))
```

#### Final Aggregated Digitec Dataset

The result of this cleaning and preprocessing workflow is the `canton_data_final` DataFrame. It contains 26 rows, one for each Swiss canton, with the key performance indicators required for further analysis (total number of sales, number of transactions, average value of orders, most ordered brand, most ordered category). This dataset is now ready to be merged with the BFS population data.

```{r}
#| label: display-final-table
#| echo: false

# Display the final, clean table
head(canton_data_final, n = 26)
```

### 4.2 BFS Datasets

```{r}

# Calculate number of transactions, average price, and median price for each bfsnr
digitec_grouped <- digitec_data_clean %>%
  group_by(bfsnr) %>%
  summarise(
    transactions = n(),
    avg_price = mean(price, na.rm = TRUE),
    median_price = median(price, na.rm = TRUE),
    .groups = "drop"
  )
```

## 5. Canton Dataset

### 5.1. Canton Level Dataset Merge

```{r}

# Merge on canton code
canton_merged_data <- left_join(canton_data_final, cleaned_canton_population, by = "canton")

# View merged dataset
canton_merged_data
```

### 5.2. Canton Level Post-merge Aggregation

To enable meaningful comparisons across cantons we add derived indicators. These indicators normalize raw counts into ratios and per-capita metrics, making them suitable for correlation analysis, segmentation, and strategic targeting.

The metrics are the following:

1.  Gender ratios
2.  Citizenship ratios
3.  Age structure
4.  Marital status

```{r}

 final_canton_data <- canton_merged_data %>% mutate( 
  
  # Economic indicators
  sales_per_capita = total_sales / total,
  transactions_per_capita = num_transactions / total,
  
  # Gender ratios
  share_male = male_population / total,

  # Citizenship ratios
  share_foreign = foreigner_citizenship / total,
  
  # Age structure
  share_age_0_19 = `age_from_0-19` / total,
  share_age_20_64 = `age_from_20-64` / total,
  share_age_65_plus = age_from_65_and_over / total,
  
  # Marital status
  share_single = single_status / total,
  share_married = married_status / total,
)
 
# Display final merged data with computed values (note: removal of original values.)
final_canton_data
```

## 6. Municipality Dataset

### 6.1 Municipality Level Dataset Merge

```{r}

# Merge age municipality dataset on bfsnr code
municipality_merged_data <- digitec_grouped %>%
  left_join(municipal_population_age_clean, by = "bfsnr")

# Merge gender municipality dataset on bfsnr code
municipality_merged_data <- municipality_merged_data %>%
  left_join(municipal_population_gender_clean, by = "bfsnr")

# View merged dataset
municipality_merged_data
```

### 6.2 Municipality Level Post-merge Aggregation

The metrics for the municipalities are the following:

1.  Gender ratios
2.  Citizenship ratios
3.  Age structure

```{r}
final_municipality_data <- municipality_merged_data %>%
  rename(
    male_population   = mann_all,
    female_population = frau_all,
    total_population  = total_all,
    swiss_citizenship    = total_swiss,
    foreigner_citizenship = total_foreigner
  ) %>%
  mutate(
    total_pop = coalesce(total_population, total),

    # gender + citizenship
    share_male    = male_population   / total_pop,
    share_foreign = foreigner_citizenship / total_pop,

    # new 10-year age shares
    share_age_10_19  = age_10_19  / total_pop,
    share_age_20_29  = age_20_29  / total_pop,
    share_age_30_39  = age_30_39  / total_pop,
    share_age_40_49  = age_40_49  / total_pop,
    share_age_50_59  = age_50_59  / total_pop,
    share_age_60_69  = age_60_69  / total_pop,
    share_age_70_79  = age_70_79  / total_pop,
    share_age_80_plus = age_80_plus / total_pop
  )

```

## 7. Data visualization

We first want to fin out which categories have the widest distribution:

```{r}
#| label: category-price-spread
#| warning: false
#| message: false

# --- 1) Compute spread + transaction count per category ----------------
spread_stats <- digitec_data_clean |>
  group_by(category) |>
  summarise(
    n_obs        = n(),
    median_price = median(price, na.rm = TRUE),
    iqr_price    = IQR(price, na.rm = TRUE),
    sd_price     = sd(price, na.rm = TRUE),
    .groups = "drop"
  )


top_n_by_transactions <- 30  # first filter: keep 30 categories with most transactions
top_n_by_spread       <- 10  # second filter: from those, keep 10 with widest IQR

# --- 2) First filter by transactions, then by widest distribution ------
# step 1: keep categories with most transactions
spread_stats_tx <- spread_stats |>
  arrange(desc(n_obs)) |>
  slice_head(n = top_n_by_transactions)

# step 2: from these, keep the ones with the widest price distribution
spread_stats_final <- spread_stats_tx |>
  arrange(desc(iqr_price)) |>
  slice_head(n = top_n_by_spread)

final_cats <- spread_stats_final$category

# --- 3) Plot: categories ordered by IQR (widest at top) ---------------
plot_spread <- digitec_data_clean |>
  filter(category %in% final_cats) |>
  mutate(
    category = factor(category, levels = rev(final_cats))  # first = widest IQR
  )

p_spread <- ggplot(plot_spread, aes(x = category, y = price)) +
  geom_boxplot(
    outlier.alpha = 0.15,
    fill = viridis::viridis(1, begin = 0.4, end = 0.6)
  ) +
  coord_flip() +
  labs(
    title = "Top Categories by Transactions and Price Spread",
    subtitle = paste0(
      "First filtered to ", top_n_by_transactions,
      " categories with most transactions,\nthen ",
      top_n_by_spread, " with widest price distribution (IQR); plot ordered by IQR (descending)"
    ),
    x = "Category",
    y = "Price (CHF)"
  ) +
  theme_minimal()

p_spread

# --- 4) Small table: show those categories with stats ------------------
top_spread_table <- spread_stats_final |>
  mutate(
    median_price = round(median_price, 1),
    iqr_price    = round(iqr_price, 1),
    sd_price     = round(sd_price, 1)
  )

top_spread_table




```

Lets focus on notebooks and smartphones, which are both economically relevant and exhibits two of the widest price distributions in the Digitec data.

### 7.1 Pre-visualization joins and processing

We first connect the cleaned Digitec transactions to the official municipality geometries.

```{r}
#| label: swiss-boundaries
#| warning: false
#| message: false

muni <- st_read("swissBOUNDARIES/swissBOUNDARIES3D_1_5_TLM_HOHEITSGEBIET.shp")


canton_lookup <- tibble::tibble(
  KANTONSNUM = 1:26,
  canton = c(
    "ZH","BE","LU","UR","SZ","OW","NW","GL","ZG","FR",
    "SO","BS","BL","SH","AR","AI","SG","GR","AG","TG",
    "TI","VD","VS","NE","GE","JU"
  )
)

muni_canton <- muni %>%
  left_join(canton_lookup, by = "KANTONSNUM")

canton_sf <- muni_canton %>%
  group_by(canton) %>%
  summarise(geometry = st_union(geometry), .groups = "drop")

canton_sf <- st_cast(canton_sf, "MULTIPOLYGON")

canton_map_data <- canton_sf %>%
  left_join(final_canton_data, by = "canton")

```

### 7.2 Notebooks

We aggregate all notebook transactions at the cantonal level, computing for each canton the total number of notebook transactions and the median and mean notebook price. These indicators are then joined with the official canton geometries to create choropleth maps that reveal spatial patterns in notebook demand and pricing.

```{r}
#| label: canton-notebook-agg
#| warning: false
#| message: false



# category name for notebooks 
notebook_cat <- "notebook-6"

# Aggregate notebook transactions per canton
notebook_canton <- digitec_data_clean |>
  filter(
    canton %in% valid_swiss_cantons,
    category == notebook_cat,
  ) |>
  group_by(canton) |>
  summarise(
    n_notebook          = n(),
    median_price_nb     = median(price, na.rm = TRUE),
    mean_price_nb       = mean(price,   na.rm = TRUE),
    .groups = "drop"
  )

# Join notebook stats to canton geometry
canton_notebook_sf <- canton_sf |>
  left_join(notebook_canton, by = "canton")

canton_notebook_sf <- canton_sf |>
  left_join(notebook_canton, by = "canton") |>
  mutate(
    # labels for transactions
    label_nb_trans = ifelse(
      is.na(n_notebook),
      canton,
      paste0(canton, " (", scales::comma(n_notebook), ")")
    ),
    # labels for prices
    label_nb_price = ifelse(
      is.na(median_price_nb),
      canton,
      paste0(canton, " (", round(median_price_nb), ")")
    )
  )



```

**Map 1 – Notebook transactions per canton**

```{r}
#| label: map-notebook-transactions-canton
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 10

# Drop NA canton (Liechtenstein)
canton_notebook_sf_plot <- canton_notebook_sf |>
  filter(!is.na(canton))

# Centroids for labels
canton_notebook_centroids <- canton_notebook_sf_plot |>
  st_centroid()

p_nb_trans_canton <- ggplot() +
  geom_sf(
    data   = canton_notebook_sf_plot,
    aes(fill = n_notebook),
    colour   = "grey30",
    linewidth = 0.2
  ) +
  geom_sf_label(
    data  = canton_notebook_centroids,
    aes(label = label_nb_trans),
    size          = 6,
    label.size    = 0.1,
    label.padding = unit(0.1, "lines"),
    fill          = "white",
    alpha         = 0.7,
    colour        = "black"
  ) +
  scale_fill_viridis_c(
    option   = "C",
    na.value = "white",
    labels   = scales::comma,
    name     = "Notebook\nTransactions"
  ) +
  labs(
    title = "Notebook Transaction Numbers",
    x = NULL, y = NULL
  ) +
  theme_minimal()

p_nb_trans_canton









```

The map shows that notebook demand is highly concentrated in the big, urban cantons, while several small / peripheral cantons have very few or no notebook transactions in our period, underlining how strongly notebook sales follow population density and economic activity.

**Map 2 – Median notebook price per canton**

```{r}
#| label: map-notebook-price-canton
#| warning: false
#| message: false
#| fig-width: 10
#| fig-height: 10

library(sf)
library(dplyr)
library(ggplot2)
library(scales)

# Drop NA canton (Liechtenstein) 
canton_notebook_sf_plot <- canton_notebook_sf |>
  filter(!is.na(canton))

# Centroids for notebook labels
canton_notebook_centroids <- canton_notebook_sf_plot |>
  st_centroid()

p_nb_price_canton <- ggplot() +
  geom_sf(
    data   = canton_notebook_sf_plot,
    aes(fill = median_price_nb),
    colour   = "grey30",
    linewidth = 0.2
  ) +
  geom_sf_label(
    data  = canton_notebook_centroids,
    aes(label = label_nb_price),
    size          = 6,
    label.size    = 0.1,
    label.padding = unit(0.1, "lines"),
    fill          = "white",
    alpha         = 0.7,
    colour        = "black"
  ) +
  scale_fill_viridis_c(
    option   = "C",
    na.value = "white",
    labels   = scales::label_comma(suffix = " CHF"),
    name     = "Median\nNotebook Price"
  ) +
  labs(
    title = "Median Notebook Price per Canton",
    x = NULL, y = NULL
  ) +
  theme_minimal()

p_nb_price_canton



```

With the Price map we can see that the high transaction number doesn't mean high median prices, but how can we use demographics to find the reason behind these price distributions?

Lets see if that has anything to do with age distribution

```{r}
#| label: using-municipality-ages-for-cantons
#| code-fold: true 
# Create a simple lookup bfsnr -> canton (no geometry)
municipality_canton_lookup <- muni_canton |>
  st_drop_geometry() |>
  select(bfsnr = BFS_NUMMER, canton)

# Aggregate municipality age counts to canton level
canton_age_from_muni <- final_municipality_data |>
  left_join(municipality_canton_lookup, by = "bfsnr") |>
  group_by(canton) |>
  summarise(
    total_pop = sum(total_pop, na.rm = TRUE),

    age_10_19 = sum(age_10_19, na.rm = TRUE),
    age_20_29 = sum(age_20_29, na.rm = TRUE),
    age_30_39 = sum(age_30_39, na.rm = TRUE),
    age_40_49 = sum(age_40_49, na.rm = TRUE),
    age_50_59 = sum(age_50_59, na.rm = TRUE),
    age_60_69 = sum(age_60_69, na.rm = TRUE),
    age_70_79 = sum(age_70_79, na.rm = TRUE),
    age_80_plus = sum(age_80_plus, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    share_age_10_19  = age_10_19  / total_pop,
    share_age_20_29  = age_20_29  / total_pop,
    share_age_30_39  = age_30_39  / total_pop,
    share_age_40_49  = age_40_49  / total_pop,
    share_age_50_59  = age_50_59  / total_pop,
    share_age_60_69  = age_60_69  / total_pop,
    share_age_70_79  = age_70_79  / total_pop,
    share_age_80_plus = age_80_plus / total_pop
  )

```

```{r}
#| label: nb-age-price-groups
#| warning: false
#| message: false
#| fig-height: 6


# Start from notebook stats and join canton ages derived from municipalities
nb_demo <- notebook_canton |>
  left_join(canton_age_from_muni, by = "canton") |>
  # add foreigner share from final_canton_data
  left_join(
    final_canton_data |> dplyr::select(canton, share_foreign),
    by = "canton"
  )

# (Optional) keep only cantons with at least 20 notebook transactions
# nb_demo <- nb_demo |> filter(n_notebook >= 20)

# Define a price threshold (median of canton median prices)
price_threshold <- median(nb_demo$median_price_nb, na.rm = TRUE)

# Group cantons into high-price vs low-price
nb_demo_groups <- nb_demo |>
  mutate(
    price_group = ifelse(
      median_price_nb >= price_threshold,
      "Higher median notebook price",
      "Lower median notebook price"
    )
  )

# Compute average age shares by price group using the finer bands
age_compare_avg <- nb_demo_groups |>
  summarise(
    share_age_10_19  = mean(share_age_10_19,  na.rm = TRUE),
    share_age_20_29  = mean(share_age_20_29,  na.rm = TRUE),
    share_age_30_39  = mean(share_age_30_39,  na.rm = TRUE),
    share_age_40_49  = mean(share_age_40_49,  na.rm = TRUE),
    share_age_50_59  = mean(share_age_50_59,  na.rm = TRUE),
    .by = price_group
  ) |>
  pivot_longer(
    cols = starts_with("share_age_"),
    names_to = "age_group",
    values_to = "share"
  ) |>
  mutate(
    age_group = factor(
      age_group,
      levels = c("share_age_10_19", "share_age_20_29", "share_age_30_39","share_age_40_49", "share_age_50_59" ),
      labels  = c("10–19", "20–29", "30–39", "40–49", "50–59")
    )
  )

ggplot(age_compare_avg,
       aes(x = age_group, y = share, fill = price_group)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_viridis_d(
    option = "C",
    begin  = 0.2,
    end    = 0.8,
    name   = NULL
  ) +
  labs(
    title = "Average age structure (municipality-based) by notebook price group",
    x     = "Age group",
    y     = "Average share of population"
  ) +
  theme_minimal()


```

The age groups show that those cantons who like more expensive notebooks have a higher share of ages between 30 and 50.

Lets also have a look at the foreigner ratio:

```{r}
#| label: nb-foreign-price-groups
#| warning: false
#| message: false
#| fig-height: 5



# Reuse nb_demo_groups from above

foreign_compare_avg <- nb_demo_groups |>
  summarise(
    avg_share_foreign = mean(share_foreign, na.rm = TRUE),
    .by = price_group
  )

ggplot(foreign_compare_avg,
       aes(x = price_group, y = avg_share_foreign, fill = price_group)) +
  geom_col() +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_viridis_d(
    option = "C",
    begin  = 0.2,
    end    = 0.8,
    guide  = "none"
  ) +
  labs(
    title    = "Average foreign resident share by notebook price group",
    x        = NULL,
    y        = "Average share of foreign residents"
  ) +
  theme_minimal()


```

Here we can actually see that the foreign ratio had an impact on the median price, which is higher when the foreign ratio is higher. This does not prove causality, but it does suggest a meaningful association: regions with a larger foreign population tend to purchase more expensive notebooks on average.


If we now distinguish between the foreigner and Swiss people for age groups we see that foreigner share is generally younger then the Swiss share.
```{r}
# age + foreign share per canton (Swiss cantons only)
age_foreign_demo <- canton_age_from_muni |>
  left_join(
    final_canton_data |> select(canton, share_foreign),
    by = "canton"
  ) |>
  filter(!is.na(canton), !is.na(share_foreign))   # <- remove FL / NAs


# split cantons into high vs low foreign share
foreign_threshold <- median(age_foreign_demo$share_foreign, na.rm = TRUE)

age_foreign_groups <- age_foreign_demo |>
  mutate(
    foreign_group = if_else(
      share_foreign >= foreign_threshold,
      "Higher foreign share",
      "Lower foreign share"
    )
  )

# compare age structure between those two groups
age_foreign_compare <- age_foreign_groups |>
  summarise(
    share_age_10_19 = mean(share_age_10_19, na.rm = TRUE),
    share_age_20_29 = mean(share_age_20_29, na.rm = TRUE),
    share_age_30_39 = mean(share_age_30_39, na.rm = TRUE),
    share_age_40_49  = mean(share_age_40_49,  na.rm = TRUE),
    share_age_50_59  = mean(share_age_50_59,  na.rm = TRUE),
    .by = foreign_group
  ) |>
  tidyr::pivot_longer(
    cols = starts_with("share_age_"),
    names_to = "age_group",
    values_to = "share"
  ) |>
  mutate(
    age_group = factor(
      age_group,
      levels = c("share_age_10_19", "share_age_20_29", "share_age_30_39","share_age_40_49", "share_age_50_59" ),
      labels  = c("10–19", "20–29", "30–39", "40–49", "50–59")
    )
  )

ggplot(age_foreign_compare,
       aes(x = age_group, y = share, fill = foreign_group)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_viridis_d(option = "C", begin = 0.2, end = 0.8, name = NULL) +
  labs(
    title = "Age structure in cantons with high vs low foreign share",
    x = "Age group",
    y = "Average share of population"
  ) +
  theme_minimal()

```
That explains why cantons with higher share of foreigners generally buy more expensive notebooks. One reason could be because younger professionals and students are more willing to spend on performance laptops. They also care about brand, specs, GPU, portability, RPM/SSD, refresh rate and are more aware of premium tech trends
```{r}
#| label: canton-smartphone-agg
#| warning: false
#| message: false
#| echo: false



# Category name for smartphones
smartphone_cat <- "smartphone-24"   

# Aggregate smartphone transactions per canton
smartphone_canton <- digitec_data_clean |>
  filter(
    canton %in% valid_swiss_cantons,
    category == smartphone_cat
  ) |>
  group_by(canton) |>
  summarise(
    n_smartphone      = n(),
    median_price_sp   = median(price, na.rm = TRUE),
    mean_price_sp     = mean(price,   na.rm = TRUE),
    .groups = "drop"
  )

# Join smartphone stats to canton geometry
canton_smartphone_sf <- canton_sf |>
  left_join(smartphone_canton, by = "canton") |>
  mutate(
    # labels for transactions: 
    label_sp_trans = ifelse(
      is.na(n_smartphone),
      canton,
      paste0(canton, " (", scales::comma(n_smartphone), ")")
    ),
    # labels for prices
    label_sp_price = ifelse(
      is.na(median_price_sp),
      canton,
      paste0(canton, " (", round(median_price_sp), ")")
    )
  )

# Join smartphone stats with canton demographics
sp_demo <- smartphone_canton |>
  left_join(final_canton_data, by = "canton")

# Define price threshold (median of canton median smartphone prices)
price_threshold_sp <- median(sp_demo$median_price_sp, na.rm = TRUE)

# Create price groups 
sp_demo_groups <- sp_demo |>
  mutate(
    price_group = if_else(
      median_price_sp >= price_threshold_sp,
      "Higher median smartphone price",
      "Lower median smartphone price"
    )
  )
```

### 7.3 Smartphone

We also did the same thing for smartphones and although the highest prices are now originating from other cantons and the foreigner share ratio patterns are not as strong they are still similar to notebook results.

```{r}
#| label: map-smartphone-price-canton
#| warning: false
#| message: false
#| echo: false
#| fig-width: 10
#| fig-height: 10


# drop NA canton (Liechtenstein)
canton_smartphone_sf_plot <- canton_smartphone_sf |>
  filter(!is.na(canton))

# centroids for label positions
canton_smartphone_centroids <- canton_smartphone_sf_plot |>
  st_centroid()

p_sp_price_canton <- ggplot() +
  geom_sf(
    data   = canton_smartphone_sf_plot,
    aes(fill = median_price_sp),
    colour   = "grey30",
    linewidth = 0.2
  ) +
  geom_sf_label(
    data  = canton_smartphone_centroids,
    aes(label = label_sp_price),
    size        = 6,
    label.size  = 0.1,
    label.padding = unit(0.1, "lines"),
    fill        = "white",
    alpha       = 0.7,
    colour      = "black"
  ) +
  scale_fill_viridis_c(
    option   = "C",
    na.value = "white",
    labels   = scales::label_comma(suffix = " CHF"),
    name     = "Median-\nSmartphone Price"
  ) +
  labs(
    title    = "Median Smartphone Price per Canton",
    x = NULL, y = NULL
  ) +
  theme_minimal()

p_sp_price_canton

```

```{r}
#| label: sp-foreign-price-groups
#| warning: false
#| message: false
#| fig-height: 5
#| echo: false



foreign_compare_sp <- sp_demo_groups |>
  summarise(
    avg_share_foreign = mean(share_foreign, na.rm = TRUE),
    .by = price_group
  )

ggplot(foreign_compare_sp,
       aes(x = price_group, y = avg_share_foreign, fill = price_group)) +
  geom_col() +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_viridis_d(option = "C", begin = 0.2, end = 0.8, guide = "none") +
  labs(
    title    = "Average foreign resident share by smartphone price group",
    x        = NULL,
    y        = "Average share of foreign residents"
  ) +
  theme_minimal()

```

```{r}
#| label: smartphone-age
#| echo: false
# --- 1) Build smartphone demo table: prices + ages + foreign share ----

sp_demo <- smartphone_canton |>
  # add age structure (10–19, 20–29, …) aggregated from municipalities
  left_join(canton_age_from_muni, by = "canton") |>
  # add foreigner share from canton-level data
  left_join(
    final_canton_data |> dplyr::select(canton, share_foreign),
    by = "canton"
  )

# --- 2) Define price threshold for smartphones ------------------------

price_threshold_sp <- median(sp_demo$median_price_sp, na.rm = TRUE)

# --- 3) Create smartphone price groups (this is sp_demo_groups) -------

sp_demo_groups <- sp_demo |>
  mutate(
    price_group = if_else(
      median_price_sp >= price_threshold_sp,
      "Higher median smartphone price",
      "Lower median smartphone price"
    )
  )




# sp_demo_groups has: canton, median_price_sp, share_foreign, age shares, price_group

age_compare_sp <- sp_demo_groups |>
  summarise(
    share_age_10_19 = mean(share_age_10_19, na.rm = TRUE),
    share_age_20_29 = mean(share_age_20_29, na.rm = TRUE),
    share_age_30_39 = mean(share_age_30_39, na.rm = TRUE),
    share_age_40_49 = mean(share_age_40_49, na.rm = TRUE),
    share_age_50_59 = mean(share_age_50_59, na.rm = TRUE),
    .by = price_group
  ) |>
  tidyr::pivot_longer(
    cols = starts_with("share_age_"),
    names_to = "age_group",
    values_to = "share"
  ) |>
  mutate(
    age_group = factor(
      age_group,
      levels = c("share_age_10_19","share_age_20_29","share_age_30_39",
                 "share_age_40_49","share_age_50_59"),
      labels  = c("10–19","20–29","30–39","40–49","50–59")
    )
  )

ggplot(age_compare_sp,
       aes(x = age_group, y = share, fill = price_group)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_viridis_d(option = "C", begin = 0.2, end = 0.8, name = NULL) +
  labs(
    title = "Age structure by smartphone price group",
    x     = "Age group",
    y     = "Average share of population"
  ) +
  theme_minimal()

```



### 7.4 Recommendations

We classify each canton based on notebook median price, share of foreigner and age structure to be able to give canton based recommendations:

```{r}
# segmentation base dataset
seg_data <- nb_demo |> 
  select(
    canton,
    median_price_nb,
    share_foreign,
    share_age_20_29,   
    share_age_30_39
  )


foreign_thr <- median(seg_data$share_foreign, na.rm = TRUE)
price_thr   <- median(seg_data$median_price_nb, na.rm = TRUE)
age_thr     <- median(seg_data$share_age_20_29 + seg_data$share_age_30_39, na.rm = TRUE)



segmented_cantons <- seg_data |>
  mutate(
    is_high_foreign = share_foreign >= foreign_thr,
    is_high_price   = median_price_nb >= price_thr,
    is_younger      = (share_age_20_29 + share_age_30_39) >= age_thr,
    
    canton_segment = case_when(
      # PREMIUM segment
      is_high_foreign & is_high_price & is_younger ~ "Premium Tech Segment",
      
      # VALUE segment
      !is_high_foreign & !is_high_price & !is_younger ~ "Value Segment",
      
      # Everyone else
      TRUE ~ "Mid-Range Segment"
    )
  )


segmented_cantons |> 
  arrange(canton_segment, desc(median_price_nb))

```
We do the same for smartphones as well:
```{r}
# --- 1) Build smartphone demo table (prices + ages + foreign share) ----

sp_demo <- smartphone_canton |>
  left_join(canton_age_from_muni, by = "canton") |>
  left_join(
    final_canton_data |> dplyr::select(canton, share_foreign),
    by = "canton"
  )

# --- 2) Segmentation base dataset for smartphones ----------------------

seg_data_sp <- sp_demo |>
  select(
    canton,
    median_price_sp,
    share_foreign,
    share_age_20_29,
    share_age_30_39
  )

# thresholds (what counts as "high")
foreign_thr_sp <- median(seg_data_sp$share_foreign,     na.rm = TRUE)
price_thr_sp   <- median(seg_data_sp$median_price_sp,   na.rm = TRUE)
age_thr_sp     <- median(seg_data_sp$share_age_20_29 + 
                           seg_data_sp$share_age_30_39, na.rm = TRUE)

# --- 3) Segment cantons for smartphones -------------------------------

segmented_cantons_sp <- seg_data_sp |>
  mutate(
    is_high_foreign = share_foreign >= foreign_thr_sp,
    is_high_price   = median_price_sp >= price_thr_sp,
    is_younger      = (share_age_20_29 + share_age_30_39) >= age_thr_sp,
    
    canton_segment_sp = case_when(
      is_high_foreign & is_high_price & is_younger ~ "Premium Tech Segment",
      !is_high_foreign & !is_high_price & !is_younger ~ "Value Segment",
      TRUE ~ "Mid-Range Segment"
    )
  )

# inspect result
segmented_cantons_sp |>
  arrange(canton_segment_sp, desc(median_price_sp))

```



```{r}
#| label: plot-seg-map
#| warning: false
#| message: false
#| fig-height: 8
#| fig-width: 9

# join notebook segments onto canton geometry
canton_sf_segmented <- canton_sf |>
  dplyr::left_join(
    segmented_cantons |> dplyr::select(canton, canton_segment),
    by = "canton"
  )

# centroids for labels
canton_sf_centroids <- canton_sf_segmented |>
  sf::st_centroid()
canton_sf_seg_plot <- canton_sf_segmented |>
  dplyr::filter(!is.na(canton))

canton_sf_centroids_plot <- canton_sf_centroids |>
  dplyr::filter(!is.na(canton))

ggplot() +
  geom_sf(
    data   = canton_sf_seg_plot,
    aes(fill = canton_segment),
    colour   = "grey30",
    linewidth = 0.2
  ) +
  geom_sf_label(
    data  = canton_sf_centroids_plot,
    aes(label = canton),
    size        = 6,
    label.size  = 0.1,
    label.padding = unit(0.1, "lines"),
    fill        = "white",
    alpha       = 0.7,
    colour      = "black"
  ) +
  scale_fill_viridis_d(
    option       = "C",
    begin        = 0.1,
    end          = 0.9,
    name         = "Segment",
    na.translate = FALSE   # don't show NA in legend
  ) +
  labs(
    title = "Canton Segments for notebook Digitec Marketing Strategy",
    x = NULL, y = NULL
  ) +
  theme_minimal()

```

```{r}
#| label: map-smartphone-segments
#| warning: false
#| message: false
#| fig-height: 8
#| fig-width: 9


# 1) Join smartphone segments to canton geometry
canton_sf_segmented_sp <- canton_sf |>
  left_join(
    segmented_cantons_sp |>
      select(canton, canton_segment_sp),
    by = "canton"
  )

# 2) Drop NA canton (Liechtenstein)
canton_sf_seg_sp_plot <- canton_sf_segmented_sp |>
  filter(!is.na(canton))

# 3) Centroids for labels
canton_sf_centroids_sp_plot <- canton_sf_seg_sp_plot |>
  st_centroid()

# 4) Plot smartphone segments
ggplot() +
  geom_sf(
    data   = canton_sf_seg_sp_plot,
    aes(fill = canton_segment_sp),
    colour   = "grey30",
    linewidth = 0.2
  ) +
  geom_sf_label(
    data  = canton_sf_centroids_sp_plot,
    aes(label = canton),
    size          = 6,
    label.size    = 0.1,
    label.padding = unit(0.1, "lines"),
    fill          = "white",
    alpha         = 0.7,
    colour        = "black"
  ) +
  scale_fill_viridis_d(
    option       = "C",
    begin        = 0.1,
    end          = 0.9,
    name         = "Smartphone segment",
    na.translate = FALSE
  ) +
  labs(
    title = "Canton Segments for Smartphone Marketing Strategy",
    x = NULL, y = NULL
  ) +
  theme_minimal()

```





**Now our recommendations: **

| Segment          | Description                                   | Recommended Focus                           |
|------------------|-----------------------------------------------|----------------------------------------------|
| **Premium Tech** (pink) | Young, international, higher spending regions. | Premium notebooks & smartphones; gaming/creator devices; multilingual, performance-focused ads. |
| **Mid-Range** (purple)  | Mixed demographics with average spending. | Mid-range laptops and phones; “value for money” messaging; accessory bundles. |
| **Value** (yellow)      | Older, more price-sensitive regions.     | Affordable, easy-to-use devices; home-office sets; reliability & service messaging. |



### 7.5 Gaming sales across Switzerland

**Why We Chose to Work With Console-Related Products?**

After visualizing Notebook category, which is the most popular and highest-traffic category on Digitec, we wanted to explore something different. Console-related products represent a less dominant category, which makes them interesting from an analytical perspective. Even though they are not as popular overall, there are specific customer groups who are highly interested in gaming products (such as us) and the gaming industry itself is very large and continues to grow.

By examining a category that doesn’t perform as strongly, we aimed to identify patterns, gaps, and potential areas for improvement.


```{r}
#| label: canton-kpi-prep



## 0.1 Cheap / Expensive segments

price_q <- quantile(data_final_swiss$price, probs = c(0.25, 0.75), na.rm = TRUE)
cheap_thr <- price_q[1]
exp_thr   <- price_q[2]

data_final_swiss <- data_final_swiss %>%
  mutate(
    price_segment = case_when(
      price <= cheap_thr ~ "Cheap",
      price >= exp_thr   ~ "Expensive",
      TRUE               ~ "Mid"
    )
  )

canton_segments <- data_final_swiss %>%
  count(canton, price_segment) %>%
  group_by(canton) %>%
  mutate(share = n / sum(n)) %>%
  ungroup() %>%
  pivot_wider(
    names_from  = price_segment,
    values_from = share,
    values_fill = 0
  ) %>%
  rename(
    cheap_share     = Cheap,
    mid_share       = Mid,
    expensive_share = Expensive
  )

final_canton_data <- final_canton_data %>%
  left_join(canton_segments, by = "canton")

data_final_swiss <- data_final_swiss %>%
  mutate(
    product_group = case_when(
      str_detect(category, "smartphone")                    ~ "Smartphone",
      str_detect(category, "gaming")                        ~ "Gaming",
      str_detect(category, "sportuhr-smartwatch")           ~ "Smartwatch",
      str_detect(category, "einbaukuehlschrank|foehn")      ~ "Home/Appliance",
      TRUE                                                  ~ "Other"
    )
  )

group_shares <- data_final_swiss %>%
  group_by(canton, product_group) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(canton) %>%
  mutate(share = n / sum(n)) %>%
  ungroup() %>%
  select(canton, product_group, share) %>%
  pivot_wider(
    names_from  = product_group,
    values_from = share,
    values_fill = 0
  ) %>%
  rename(
    smartphone_share = Smartphone,
    gaming_share     = Gaming,
    smartwatch_share = Smartwatch,
    home_share       = `Home/Appliance`
  )

final_canton_data <- final_canton_data %>%
  left_join(group_shares, by = "canton")

canton_map_data <- canton_sf %>%
  left_join(final_canton_data, by = "canton")
```

```{r}


gaming_data <- data_final_swiss %>%
  filter(
    brandName == "Nintendo" |
    (brandName == "Ugreen" &
       str_detect(fullProductName, regex("switch|nintendo|joy[- ]?con", ignore_case = TRUE))) |
    str_detect(fullProductName, regex("playstation|ps4|ps5", ignore_case = TRUE)) |
    str_detect(fullProductName, regex("xbox", ignore_case = TRUE))
  ) %>%
  mutate(
    platform = case_when(
      brandName == "Nintendo" ~ "Nintendo",
      brandName == "Ugreen" &
        str_detect(fullProductName, regex("switch|nintendo|joy[- ]?con", ignore_case = TRUE)) ~ "Nintendo",
      str_detect(fullProductName, regex("playstation|ps4|ps5", ignore_case = TRUE)) ~ "PlayStation",
      str_detect(fullProductName, regex("xbox", ignore_case = TRUE)) ~ "Xbox",
      TRUE ~ "Other"
    )
  )
```

```{r}

gaming_canton <- gaming_data %>%
  group_by(canton) %>%
  summarise(
    gaming_sales        = sum(price, na.rm = TRUE),
    gaming_transactions = n(),
    avg_gaming_price    = mean(price, na.rm = TRUE),
    .groups = "drop"
  )

canton_map_data <- canton_map_data %>%
  filter(!is.na(canton)) %>%        
  select(-any_of(c("gaming_sales", 
                   "gaming_transactions", 
                   "avg_gaming_price", 
                   "tooltip_gaming",
                   "gaming_share_total"))) %>%  
  left_join(gaming_canton, by = "canton") %>%
  mutate(
    gaming_sales        = coalesce(gaming_sales, 0),
    gaming_transactions = coalesce(gaming_transactions, 0),
    avg_gaming_price    = coalesce(avg_gaming_price, 0),
    gaming_share_total  = if_else(total_sales > 0, 
                                  gaming_sales / total_sales, 
                                  0)
  ) %>%
  mutate(
    tooltip_gaming = paste0(
      "<b>Canton: </b>", canton,
      "<br><b>Gaming sales:</b> ", scales::comma(round(gaming_sales, 0)), " CHF",
      "<br><b>Transactions:</b> ", gaming_transactions,
      "<br><b>Avg basket:</b> ", round(avg_gaming_price, 1), " CHF",
      "<br><b>Gaming share:</b> ",
      scales::percent(gaming_share_total, accuracy = 0.1)
    )
  )

```






**Where are consoles the most popular?**

To understand regional differences in gaming demand, we first aggregated Nintendo, PlayStation and Xbox transactions to the canton level. We then merged these data with demographic and geographic information from the Federal Statistical Office.

The interactive map below visualizes total gaming sales by canton. Hovering over a canton displays key metrics, including total gaming sales, number of transactions, average gaming basket size, and the canton’s gaming share.

```{r}

p_gaming_sales <- ggplot(canton_map_data) +
  geom_sf(
    aes(
      fill = gaming_sales,
      text = tooltip_gaming
    ),
    color = "grey20",
    linewidth = 0.3
  ) +
  scale_fill_viridis_c(
    option = "C",
    name = "Gaming Sales [CHF]",
    labels = scales::label_comma()
  ) +
  labs(
    title = "Gaming Sales by Canton",
    subtitle = "Nintendo, PlayStation, Xbox & Ugreen (Switch)"
  ) +
  theme_minimal()

ggplotly(p_gaming_sales, tooltip = "text")

```

Gaming demand in Switzerland is heavily concentrated in urban cantons. Zurich is the single most important market for Digitec’s gaming category, followed by Bern, Aargau, Vaud, St. Gallen, and Geneva. These cantons not only drive the majority of gaming sales but also include the highest-density gamer populations based on gaming share percentage.

Cantons with zero or low gaming activity indicate rural areas or small populations where outreach will not significantly impact total sales. So, if Digitec wants to maximize return on advertising spend, focusing on the six large cantons gives access to the largest clusters of active gamers.

```{r}
top_cantons <- canton_map_data %>%
  distinct(canton, .keep_all = TRUE) %>%
  arrange(desc(gaming_sales)) %>%
  slice_head(n = 6) %>%
  pull(canton)

top_cantons

```

```{r}
gaming_top <- gaming_data %>%
  filter(canton %in% top_cantons)

platform_canton <- gaming_top %>%
  group_by(canton, platform) %>%
  summarise(
    sales = sum(price, na.rm = TRUE),
    transactions = n(),
    avg_price = mean(price, na.rm = TRUE),
    .groups = "drop"
  )

```

We next examined which gaming platforms dominate in the six strongest cantons.

```{r}

ggplot(platform_canton, aes(x = canton, y = sales, fill = platform)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = comma) +
  scale_fill_viridis_d(option = "C") +
  labs(
    title = "Platform Breakdown in Top Gaming Cantons",
    x = "Canton",
    y = "Gaming Sales [CHF]",
    fill = "Platform"
  ) +
  theme_minimal(base_size = 14)
```

In **five out of six** of these top cantons (AG, BE, GE, SG, ZH), PlayStation generates the highest sales. Xbox is consistently the smallest of the three but still meaningful everywhere, especially in Zurich and Aargau. **Vaud (VD) is the exception where Nintendo dominates**.

In short, Digitec should push PlayStation heavily in Zürich and the German-speaking top cantons where it dominates, while tailoring Vaud with more Switch-focused offers.

We found it very interesting that Vaud (VD) is the only top canton where Nintendo beats PlayStation in sales. A Nintendo-specific regional campaign will perform better here than a broad PlayStation related campaigns. This makes it Digitec’s strongest opportunity for a targeted Switch-focused campaign. By breaking down Nintendo sales inside Vaud’s municipalities, we can identify exactly where the most engaged family and casual gamers live.

So the *right question* now is:

**Why is Vaud different? Who actually buys all this Nintendo products?**

To understand why Nintendo is so strong in Vaud, we analyzed how gaming purchasing relates to the canton’s age structure.

```{r}
vaud_demo <- canton_map_data %>%
  filter(canton == "VD") %>%
  slice(1) %>%   
  mutate(
    share_age_0_19 = `age_from_0-19` / total,
    share_age_20_64 = `age_from_20-64` / total,
    share_age_65_plus = age_from_65_and_over / total
  )
vaud_nintendo <- gaming_data %>%
  filter(canton == "VD", platform == "Nintendo") %>%
  select(fullProductName, price)

total_nintendo_sales_vaud <- sum(vaud_nintendo$price, na.rm = TRUE)

vaud_nintendo_age <- tibble(
  age_group = c("0_19", "20_64", "65_plus"),
  sales = c(
    total_nintendo_sales_vaud * vaud_demo$share_age_0_19,
    total_nintendo_sales_vaud * vaud_demo$share_age_20_64,
    total_nintendo_sales_vaud * vaud_demo$share_age_65_plus
  )
)


```

```{r}
total_nintendo_sales_vaud <- sum(vaud_nintendo$price, na.rm = TRUE)

vaud_nintendo_age <- tibble(
  age_group = c("0_19", "20_64", "65_plus"),
  sales = c(
    total_nintendo_sales_vaud * vaud_demo$share_age_0_19,
    total_nintendo_sales_vaud * vaud_demo$share_age_20_64,
    total_nintendo_sales_vaud * vaud_demo$share_age_65_plus
  )
)

ggplot(vaud_nintendo_age,
       aes(x = age_group, y = sales)) +
  geom_col(fill = "#FF4B4B") +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Nintendo Sales in Vaud by Age Segment (Canton Level)",
    x = "Age Group",
    y = "Nintendo Sales [CHF]"
  ) +
  theme_minimal(base_size = 14)

```

The 20–64 age segment generates the majority of Nintendo spending. Youth (0–19) contribute strongly, while seniors play only a minor role.

This means Nintendo’s strength in Vaud comes from **adult players buying premium products** or from parents purchasing consoles for children. But what do they buy? Do they buy products for themselves or their children?

```{r}
vaud_demo <- canton_map_data %>%
  filter(canton == "VD") %>%
  slice(1) %>%  
  mutate(
    share_age_20_64 = `age_from_20-64` / total
  )

vaud_nintendo <- gaming_data %>%
  filter(canton == "VD", platform == "Nintendo") %>%
  select(fullProductName, price)

vaud_nintendo_weighted <- vaud_nintendo %>%
  mutate(
    weighted_sales = price * vaud_demo$share_age_20_64
  )

nintendo_product_summary <- vaud_nintendo_weighted %>%
  group_by(fullProductName) %>%
  summarise(
    sales_20_64 = sum(weighted_sales, na.rm = TRUE),
    transactions = n(),
    .groups = "drop"
  ) %>%
  arrange(desc(sales_20_64)) %>%
  slice_head(n = 10)  


```

```{r}

#for visualization purposes
nintendo_product_summary$fullProductName <- str_wrap(nintendo_product_summary$fullProductName, width = 30)

ggplot(nintendo_product_summary,
       aes(x = reorder(fullProductName, sales_20_64),
           y = sales_20_64)) +
  geom_col(fill = "#E60012") +  # Nintendo red
  coord_flip() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Top Nintendo Products Bought by Adults (20–64) in Vaud",
    x = "Nintendo Product",
    y = "Weighted Sales [CHF]"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.text.y = element_text(size = 9), 
    plot.margin = margin(10, 20, 10, 20)     
  )


```

The top items in the chart are almost all **premium editions of switch consoles or games**. This shows that adults in Vaud are making **big Nintendo purchases**, not just small add-ons.

Vaud’s adult population buys premium Nintendo consoles and major franchises like Zelda. Zelda is a franchise primarily popular among adults, so we can say that a large portion of Nintendo customers in Vaud are adult gamers, not only parents shopping for kids as some people might think.

This means Nintendo advertising in Vaud should not focus only on kids but it must mostly target adult gamers who choose Switch for themselves.

### 7.6 Recommendations Gaming

Zurich, Bern, Aargau, St. Gallen, and Geneva all show a clear preference for PlayStation. These cantons are also the largest contributors to gaming sales overall. So, Digitec should push **PS5 bundles, controllers, PS Plus subscriptions, and headline releases** heavily in these markets.

Vaud is the only major canton where Nintendo dominates PlayStation. Adults (20–64) drive most of these sales, especially premium Switch models and Zelda titles. Campaigns here should focus on **adult Nintendo players**, premium Switch upgrades and promotions on favourite fanchises like Zelda.

# 8. Conclusions & Strategic Recommendations

Our analysis of Digitec transaction records combined with FSO demographic data confirms that a generic, nationwide marketing strategy is inefficient. The data reveals significant regional and demographic disparities that necessitate a shift towards "Smart Targeting."

We demonstrated that sales volume and sales value follow different geographic patterns and that cultural differences (the "Röstigraben") influence purchasing behavior more strongly than expected.

### Key Insights

1.  **The Volume-Value Paradox:** While sales volume (number of transactions) strongly correlates with population density (Zurich, Bern), these urban centers do not necessarily yield the highest average basket sizes. Our analysis highlights a strong correlation between the share of foreign residents and higher median prices for notebooks and smartphones. This points to high-purchasing-power clusters (e.g., in Zug or Geneva) that are significantly more likely to purchase premium products than the average consumer.

2.  **The "Gaming Divide" & The Vaud Anomaly:** While German-speaking Switzerland (Zurich, St. Gallen, Aargau) is clearly dominated by PlayStation, the Canton of Vaud exhibits a unique preference for Nintendo. Crucially, our deep-dive analysis refutes the "kids' console" stereotype: It is the 20–64 age demographic in Vaud that drives spending on premium Nintendo products (OLED models, Zelda franchise), rather than parents buying for children.

### Strategic Recommendations

Based on these findings, we recommend the following strategic actions for Digitec:

-   **Purchasing Power Segmentation (High-End Targeting):** Shift advertising budgets for premium hardware (MacBook Pro, iPhone Pro, Samsung Ultra) towards cantons with high expat densities and purchasing power (ZG, GE, BS). In traditional population hubs (ZH, BE), marketing should focus on value-for-money and product variety to maintain transaction volume.

-   **Regionally Differentiated Gaming Marketing:**

    -   **German-speaking Switzerland:** Focus on PlayStation performance, competitive gaming, and hardware bundles.
    -   **Romandy (Vaud):** Launch a dedicated Nintendo campaign specifically targeting **adult gamers**. The messaging should emphasize premium hardware and sophisticated gameplay (e.g., Zelda), moving away from child-centric imagery.

## 8.1 Campaign Concept: "Switzerland in Data"

To translate these strategic findings into customer-facing communication and demonstrate product variety, we designed four visuals following the established Galaxus campaign style ("Our range explained with data"). These are directly based on the four "catchy findings" from our analysis:

1.  **The Vaud Paradox (Gaming):** Based on our age structure analysis, highlighting that Nintendo in Vaud is primarily purchased by adults (20–64).
    -   *Headline:* "Why do almost only adults buy this game in Vaud?"
2.  **The Zug Premium (Notebooks):** Illustrates the massive gap in median basket size between Zug and Bern, driven by the purchasing power in Zug.
    -   *Headline:* "Why does Zug buy more expensive laptops than Bern?"
3.  **The Console Divide (Platforms):** Visualizes the cultural split in gaming behavior: PlayStation dominates Zurich, while Vaud prefers Nintendo.
    -   *Headline:* "Why does Zurich play PlayStation and Vaud Nintendo?"
4.  **The Geneva Connection (Smartphones):** Leverages our correlation analysis regarding foreign resident shares. Geneva (high expat share) shows a significantly higher willingness to pay for smartphones than the average.
    -   *Headline:* "Why does Geneva swipe more expensively?"

## 9. Individual contributions

| Name | Contribution |
|---------------|--------------------------------------------------------|
| Anna | Business plan analysis, data cleaning, pre-visualization processing, gaming sales vizualizations |
| Fereshteh | Business plan analysis, data cleaning, pre-visualization processing, Notebooks and smartphone vizualizations |
| Sen |  |
| Colin | Digitec Data Cleaning and Preprocessing, Basic Data Visualization, Conclusion and Commercial Recommandation |
