---
title: "Digitec Analysis"
format:
  html:
    df-print: paged
    toc: true
    code-tools: true

editor: visual
author: Fereshteh, Anna, Sen, Colin Marti (pls add your full name)
---

## 1. Business Plan Analysis

Digitec Galaxus is Switzerland’s largest online retailer for consumer electronics, operating exclusively via e-commerce. Despite nationwide reach, customer behavior likely differs across cantons due to income, urbanization, age, citizenship, and household structures.

This project explores how these demographic and regional factors influence online purchases using: - \~163,000 Digitec transaction records, and - Canton-level demographic data from the Swiss Federal Statistical Office (BFS).

The findings will guide a data-driven marketing strategy to personalize digital campaigns and increase online sales efficiency.

### 1.1. Business Problem

> **How can regional demographics be used to improve Digitec’s online marketing strategy and boost conversions across Switzerland?**

Digitec currently uses uniform national marketing, but demographic diversity suggests that: - Urban cantons might favor high-end tech, - Rural areas could prefer affordable products, - Younger, single populations may show distinct category interests.

By linking demographic patterns with online purchasing data, Digitec can **target audiences more intelligently** and tailor its digital campaigns for maximum impact.

### 1.2. Data Overview

| Source | Description | Key Variables |
|:-----------------------|:-----------------------|:-----------------------|
| **Digitec Transactions (`DigitecLive_Final.csv`)** | \~163k web-scraped online transactions from Digitec’s live purchase feed | `brandName`, `infos.Category`, `salesPrice.amountIncl`, `dateTime`, `mun_canton` |
| **BFS Population Data (`population_2023.xlsx`)** | Canton-level demographics (2023) | `Total`, `Swiss`, `Foreigner`, `Urban`, `Intermediate`, `Rural`, `Age_0_19`, `Age_20_64`, `Age_65plus`, `Single`, `Married`, `Widowed`, `Divorced` |
| **BFS Population by Age (`mns-population-by-age-2023.xlsx`)** | Detailed population structure by age and region | `Region`, `Total`, `Age_0_4`, `Age_5_9`, …, `Age_80plus` |
| **BFS Population by Gender (`mns-population-by-gender-2023.xlsx`)** | Gender composition by canton and municipality (2023) | `Region`, `Total_All`, `Mann_All`, `Frau_All`, `Swiss`, `Foreigner` |
| **Swiss Municipal Boundaries (`swissBOUNDARIES3D_1_5_TLM_HOHEITSGEBIET.shp`)** | Official Swiss municipality polygons used to map Digitec transactions geographically and join by BFS municipality number | `BFS_NUMMER` (municipality ID), `NAME` (municipality name), `KANTONSNUM` (canton ID), `geometry` (MULTIPOLYGON) |

These additional datasets allow the inclusion of **more granular demographic information** — enabling a refined analysis of age- and gender-related purchasing behavior across regions.

### 1.3. Unit of Analysis

The analysis is now conducted at the **municipality level** ($n \approx 2,200$), using the official **BFS municipality number (`mun_bfsnr`)** as the join key between datasets.

Each municipality represents one observation containing: - $sales\_per\_capita = \frac{total\_sales}{population}$\
- Average order value (CHF)\
- Age-group shares (e.g., % under 20, % over 65)\
- Gender composition (% male, % female)\
- Citizenship composition (% Swiss, % foreign)

This finer-grained approach allows for **more localized insights**, capturing demographic and purchasing differences that are hidden when aggregating by canton.

------------------------------------------------------------------------

### 1.4. Methodology

#### 1.4.1. Analytical Steps

1.  **Aggregate** Digitec transaction data by municipality (`mun_bfsnr`)\
    → Compute total sales, average price, and transaction count for each municipality.\
2.  **Merge** the municipality-level BFS datasets on population, gender, and age using the `mun_bfsnr` key.\
3.  **Compute derived demographic indicators:**
    -   Gender ratio: $\frac{Male}{Female}$\
    -   Share of foreign residents: $\frac{Foreigner}{Total}$\
    -   Age structure: $\frac{Population_{0-19}}{Total}$, $\frac{Population_{20-64}}{Total}$, $\frac{Population_{65+}}{Total}$\
    -   Sales per capita and average order value\
4.  **Explore correlations** between online sales indicators and demographics:
    -   Are younger or more international municipalities buying more expensive products?\
    -   Do gender or age structures explain variations in spending per capita?\
5.  **Visualize** geographic and demographic patterns using:
    -   Choropleth maps by municipality (sales per capita, average price, etc.)\
    -   Scatterplots linking demographics to sales variables\
    -   Heatmaps of demographic vs. economic indicators\
6.  **Identify four key insights** to inform Digitec’s new data-driven marketing campaign.

#### 1.4.2. Research Questions

| **Demographic Dimension** | **Research Question** | **Usage Example (What Digitec Could Do)** |
|------------------------|------------------------|------------------------|
| **Population Size** | Do larger cantons generate more online purchases? | Focus major campaigns and product launches in high-activity cantons, while using localized ads to increase engagement in smaller regions. |
| **Gender Composition** | Does the male/female balance influence which product categories are bought? | Promote gaming content in female-dominant cantons and highlight smartphone or smartwatch deals in male-dominant regions. |
| **Citizenship** | Do cantons with more foreign residents show different tech-buying patterns? | Use multilingual ads (DE/FR/EN) and highlight global brands such as Apple or Samsung in foreign-dense cantons. |
| **Urbanization** | Do urban regions buy more high-end devices compared to rural regions? | Promote flagships and new releases in urban cantons; emphasize budget or refurbished products in rural areas. |
| **Marital Status** | Do single vs. married populations differ in what they purchase? | Focus gaming offers in single-dominant cantons; promote smart-home and smartwatch bundles in married-dominant regions. |

### 1.5. Strategic Impact

By merging **online sales data** with **BFS demographic indicators**, Digitec can: - Move from national-level campaigns to **regionally adaptive marketing**.\
- Align ad language, pricing, and products with **local audience characteristics**.\
- Strengthen relevance and **optimize ad spend efficiency**.

## 2. Setup and Data Loading

This first code chunk sets up the R environment by loading all necessary libraries for the analysis and loading the four primary datasets. The chunk is configured with `include: false` to hide it from the final rendered report for a cleaner presentation.

```{r}

#| label: setup
#| include: false

# Load all required libraries for the project
library(tidyverse) # For data manipulation (dplyr) and visualization (ggplot2)
library(readxl)    # To read Excel files
library(lubridate) # To work with date-time values
library(stringr)   # To work with regex
library(dlookr)
library(ggrepel)
library(ggcorrplot)
library(viridis)
library(dplyr)
library(plotly)
library(sf)
library(scales)    

# --- GLOBAL DESIGN DEFINITION ---
# Save the original theme_minimal
orig_theme_minimal <- ggplot2::theme_minimal

# Override with custom settings
theme_minimal <- function(base_size = 11, base_family = "", ...) {
  orig_theme_minimal(base_size = base_size, base_family = base_family, ...) +
  theme(
    # 1. Axis text (values on ticks)
    axis.text = element_text(size = 18, color = "black"),
    
    # 2. Axis titles 
    axis.title = element_text(size = 20, color = "black", margin = margin(t=15, r=15)),
    
    # 3. Main Plot Title
    plot.title = element_text(size = 26, color = "black", margin = margin(b=15)),
    
    # 4. Plot Subtitle
    plot.subtitle = element_text(
      size = 20, 
      color = "black", 
      margin = margin(b=20)
    ),
    
    # 5. Caption 
    plot.caption = element_text(size = 14, color = "grey30", hjust = 1),
    
    # 6. Legends
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18)
  )
}

# Load digitect dataset
load("digitec.RData") # Loads the 'data' object

# Save to descriptive variable
digitec_data <- get("data")

## Load BFS data sets
canton_population <- read_excel("population_2023.xlsx")
municipal_population_gender <- read_excel("mns-population-by-gender-2023.xlsx")
municipal_population_age <- read_excel("mns-population-by-age-2023.xlsx")

```

## 3. Data Cleaning and Preprocessing

This section details the systematic process of cleaning and preparing the raw data of the required data sets. The goal is to produce a reliable, aggregated dataset at the municipal and cantonal level. The municipal data will be a combination of three datasets: "mns-population-by-age", "mns-population-by-gender", and the digitec dataset. The cantonal dataset will be a merging of the BFS demographic data and the digitec dataset.

### 3.1. Digitec Dataset

An initial inspection of the raw Digitec data revealed several necessary cleaning steps:

-   **Correcting Data Types:** The `dateTime` column was stored as a character string and needed to be converted into a proper date-time format for time-based analysis.
-   **Removing Irrelevant Columns:** Many columns were redundant (e.g., price with and without VAT), contained technical metadata, or were not relevant to the analysis (e.g., image URLs).
-   **Ensuring Data Quality:** Transactions without a price or a clear, unique geographic match (`mun_nhits != 1`) were removed to ensure the reliability of the analysis.
-   **Feature Engineering:** More useful features were extracted from the raw data, such as the day of the week from the `dateTime` column or the clean product category name from a URL string.

#### 3.1.1. Data Cleaning & Engineering

First we check the dataset columns that we think might require some filtering or deletion.

```{r}
candidates <- c(
  "cityName",
  "salesPrice.currency",
  "infos.Price",
  "infos.Currency",
  "salesPrice",
  "mun_nhits"
)

existing <- intersect(candidates, names(digitec_data))

check_var <- function(df, var) {
  x <- df[[var]]
  tibble(
    variable      = var,
    exists        = TRUE,
    n             = length(x),
    n_non_na      = sum(!is.na(x)),
    prop_na       = mean(is.na(x)),
    n_unique_non_na = dplyr::n_distinct(x, na.rm = TRUE),
    example_values  = paste(utils::head(unique(x), 3), collapse = " | ")
  )
}

profile <- map_dfr(existing, ~check_var(digitec_data, .x))
profile
```

-   `salesPrice.currency`, `infos.Currency` → `n_unique_non_na == 1` (“CHF”) we decided to remove this column.
-   `salesPrice` → `n_non_na == 0` empty column so we decided to delete it.
-   `mun_nhits` → mostly/all 1

```{r}
if ("mun_nhits" %in% names(digitec_data)) {
  cat("mun_nhits frequency (percent):\n")
  perc <- prop.table(table(digitec_data$mun_nhits, useNA = "ifany")) * 100
  print(round(perc, 2))
}
```

-   **`1`** → 153 778 or 94.35% of the rows matched exactly one municipality.
-   **`0`** → 9 194 or 5.64% of the rows had *no municipality match*.
-   **`2`** → 21 or 0.01% of the rows matched *two possible municipalities*.

We would drop the rows where `mun_nhits == 0` or `mun_nhits == 2`, then remove the `mun_nhits` column. Since we already have 95%.

-   `cityName` → unique city name inserted by customer, since we have the city and municipality takes from SWISS POST Api we decided to drop this column.
-   `infos.Price` → many NAs - 63% of the data is missing so the column is invalid.

```{r}
if (all(c("infos.Price", "salesPrice.amountIncl") %in% names(digitec_data))) {
  same_info <- digitec_data %>%
    filter(!is.na(infos.Price) & !is.na(salesPrice.amountIncl)) %>%
    summarise(
      n = n(),
      max_abs_diff = max(abs(infos.Price - salesPrice.amountIncl), na.rm = TRUE)
    )
  same_info
}
```

60 577 rows had *both* `infos.Price` and `salesPrice.amountIncl` . The value `max_abs_diff = 1021.9` means that in some rows, the difference between the two price fields is CHF 1 021.90 so prices differ a lot. We decided that this column is inconsistent.

The following steps were implemented in a single, coherent `dplyr` pipeline for efficiency and reproducibility. The result is a cleaned DataFrame named `data_clean`.

```{r}
#| label: data-cleaning-pipeline

library(dplyr)
library(tidyr)
library(lubridate)
library(forcats)

digitec_data_clean <- digitec_data %>%
  # 1) Select and rename relevant columns for better readability
  rename(
    price        = salesPrice.amountIncl,
    category_url = infos.Category,
    sales_rank_raw = infos.Sales_Rank,
    canton       = mun_canton,
    bfsnr        = mun_bfsnr
  ) %>%
  select(
    price, dateTime, brandName, fullProductName,
    socialShoppingTransactionTypeId,
    category_url, sales_rank_raw, canton, mun_nhits, bfsnr
  ) %>%
  
  # 2) Filter rows to ensure data quality
  filter(
    !is.na(price),                  # keep only transactions with a price
    !is.na(canton),                 # keep only transactions with canton info
    mun_nhits == 1,                 # keep only unique geographic matches
    socialShoppingTransactionTypeId %in% c(3, 5),  # only bought/shipped
    grepl("von", sales_rank_raw)    # keep only rows where rank contains "von"
  ) %>%
  
  # 3) Remove irrelevant or redundant columns
  select(-any_of(c(
    "cityName",
    "salesPrice.currency",
    "infos.Price",
    "infos.Currency",
    "salesPrice"
  ))) %>%
  
  # 4) Create new, useful columns (feature engineering)
  mutate(
    # timestamps
    date     = as_date(dateTime),
    dateTime = ymd_hms(dateTime),
    hour     = hour(dateTime),
    weekday  = wday(dateTime, label = TRUE, week_start = 1),
    # product category from URL
    category = basename(category_url)
  ) %>%
  
  # 5) Parse sales rank: "12 von 345" -> numeric position / total
  separate(
    sales_rank_raw,
    into = c("rank_position", "rank_total"),
    sep  = " von "
  ) %>%
  mutate(
    rank_position = as.numeric(rank_position),
    rank_total    = as.numeric(rank_total),
    rank_relative = rank_position / rank_total,
    rank_score    = 1 - rank_relative
  ) %>%
  
  # 6) Recode transaction type and convert key vars to factors
  mutate(
    transaction_type = case_when(
      socialShoppingTransactionTypeId == 3 ~ "bought",
      socialShoppingTransactionTypeId == 5 ~ "shipped"
    ),
    transaction_type = fct_relevel(transaction_type, "bought", "shipped"),
    category  = as.factor(category),
    brandName = as.factor(brandName)
  ) %>%
  select(-socialShoppingTransactionTypeId, -mun_nhits) %>%  # no longer needed
  
  # 7) Drop duplicates
  distinct() %>%
  
  # 8) Handle Inf and winsorize rank metrics in one go
  mutate(
    rank_score    = ifelse(is.infinite(rank_score),    NA_real_, rank_score),
    rank_relative = ifelse(is.infinite(rank_relative), NA_real_, rank_relative)
  ) %>%
  { 
    qs_score <- quantile(.$rank_score,    probs = c(0.025, 0.975), na.rm = TRUE)
    qs_rel   <- quantile(.$rank_relative, probs = c(0.025, 0.975), na.rm = TRUE)
    
    mutate(.,
      rank_score    = pmin(pmax(rank_score,    qs_score[1]), qs_score[2]),
      rank_relative = pmin(pmax(rank_relative, qs_rel[1]),   qs_rel[2])
    )
  } %>%
  filter(!is.na(rank_score), !is.na(rank_relative))
```

```{r}
overview(digitec_data_clean)
summary(digitec_data_clean)

```

### 3.2. BFS Datasets

The data within the three BFS is mostly clean. There are no apparent missing values or errors, just some unnecessary rows. More importantly, the datasets need to be compatible for the upcoming merge with the Digitec data.\
\
The issues that we identified were as follows:

-   **Column names:** The column names of all three datasets need to be non-capitalized and follow the snake_case format in order to align with the digitec dataset.

-   **Unwanted rows:** Remove the unwanted canton and bezirk rows with the municipality datasets and rename the column to muncipality.

-   **Canton codes:** The canton names within the municipality BFS dataset are not in the standard code format.

-   **Municipality number:** The age and gender datasets have an improper municipality notation. Later on, we'll have to merge these datasets with the digitec dataset according to the municipality number, so it's imperative that we make sure they're in the same format.

-   **Ages:** Currently ages within the age datasetare not grouped in the same way as BFS' canton dataset.

#### 3.2.1 Data Cleaning & Engineering

**Municipality Level**

Firstly, we will tackle the column names; we do this simply by running each dataset through this snake_case function:

```{r}

#function which converts input to snake_case
clean_names <- function(x) {
  x %>%
    str_to_lower() %>%
    
    # Replace spaces with underscores
    str_replace_all(" ", "_")             
}

# Rename each dataset
colnames(canton_population) <- clean_names(colnames(canton_population))
colnames(municipal_population_age) <- clean_names(colnames(municipal_population_age))
colnames(municipal_population_gender) <- clean_names(colnames(municipal_population_gender))
```

Then within the age and gender datasets we can remove the unnecessary rows, rename the column and extract the bfsnr from the municipalities.

```{r}
#| label: row-removal

# Rename first column to bfsnr like in the digitec dataset
colnames(municipal_population_age)[1] <- "bfsnr"
colnames(municipal_population_gender)[1] <- "bfsnr"

# Clean municipality age data
municipal_population_age_clean <- municipal_population_age %>%
  filter(str_detect(bfsnr, "^\\.+")) %>%  # Match rows starting with one or more dots
  mutate(
    bfsnr = str_extract(bfsnr, "\\d+"),  # Extract numeric code
    bfsnr = as.integer(bfsnr)
  )

# Clean municipality gender data
municipal_population_gender_clean <- municipal_population_gender %>%
  filter(str_detect(bfsnr, "^\\.+")) %>%
  mutate(
    bfsnr = str_extract(bfsnr, "\\d+"),
    bfsnr = as.integer(bfsnr)
  )

municipal_population_age_clean
municipal_population_gender_clean

```

Now we group the ages from 0-19, 20-64, and 65-100+ and our age and population datasets are ready for merging.

```{r}

# Collapse into age bands
municipal_population_age_clean <- municipal_population_age_clean %>%
  rowwise() %>%
  mutate(
    `age_from_0-19` = sum(c_across(`0`:`19`), na.rm = TRUE),
    `age_from_20-64` = sum(c_across(`20`:`64`), na.rm = TRUE),
    age_from_65_and_over = sum(c_across(`65`:`99`), `100_und_mehr`, na.rm = TRUE)
  ) %>%
  ungroup() %>% select(bfsnr, total, `age_from_0-19`, `age_from_20-64`, age_from_65_and_over)

municipal_population_age_clean
```

**Canton Level**

For the canton level we need to map the canton names found in the canton BFS dataset to their official abbreviations used in the transaction dataset.

```{r}
#| label: canton-code-mapping


# Define canton name-to-code mapping
canton_codes <- list(
  "Zurich" = "ZH",
  "Bern" = "BE",
  "Lucerne" = "LU",
  "Uri" = "UR",
  "Schwyz" = "SZ",
  "Obwalden" = "OW",
  "Nidwalden" = "NW",
  "Glarus" = "GL",
  "Zug" = "ZG",
  "Fribourg" = "FR",
  "Solothurn" = "SO",
  "Basel-Stadt" = "BS",
  "Basel-Landschaft" = "BL",
  "Schaffhausen" = "SH",
  "Appenzell A. Rh." = "AR",
  "Appenzell I. Rh." = "AI",
  "St. Gallen" = "SG",
  "Graubünden" = "GR",
  "Aargau" = "AG",
  "Thurgau" = "TG",
  "Ticino" = "TI",
  "Vaud" = "VD",
  "Valais" = "VS",
  "Neuchâtel" = "NE",
  "Geneva" = "GE",
  "Jura" = "JU"
)

# Convert to named vector for mapping
code_map <- setNames(unlist(canton_codes), names(canton_codes))

# Replace canton names with codes and remove non-cantons
cleaned_canton_population <- canton_population %>%
  filter(canton %in% names(canton_codes)) %>%
  mutate(canton = code_map[canton])


# Display the processed excel data
cleaned_canton_population

```

## 4. Validation and Final Aggregation

### 4.1 Digitec Dataset

An initial aggregation of the cleaned data surprisingly resulted in 29 groups instead of the 26 expected Swiss cantons. An investigation of the `canton` column revealed that the dataset also contained transactions from neighboring countries, which needed to be excluded for this analysis.

```{r}
#| label: validation-step

# A list of valid Swiss canton abbreviations
valid_swiss_cantons <- c(
  "AG", "AI", "AR", "BE", "BL", "BS", "FR", "GE", "GL", "GR", 
  "JU", "LU", "NE", "NW", "OW", "SG", "SH", "SO", "SZ", "TG", 
  "TI", "UR", "VD", "VS", "ZG", "ZH"
)

# Identify which unique codes in our data are NOT in the valid list
non_swiss_codes <- setdiff(unique(digitec_data_clean$canton), valid_swiss_cantons)

print("Found the following non-Swiss location codes:")
non_swiss_codes
```

Based on this finding, the dataset was filtered to include only transactions from the 26 Swiss cantons. Afterward, the final aggregation was performed to create the analysis-ready dataset.

```{r}
#| label: final-aggregation

# Filter the dataset to retain only transactions from Swiss cantons
data_final_swiss <- digitec_data_clean %>%
  filter(canton %in% valid_swiss_cantons)

# Perform the final aggregation at the cantonal level
canton_data_final <- data_final_swiss %>%
  group_by(canton) %>%
  summarise(
    total_sales = sum(price, na.rm = TRUE),
    num_transactions = n(),
    avg_order_value = mean(price, na.rm = TRUE),
    top_brand = names(which.max(table(brandName))),
    top_category = names(which.max(table(category)))
  ) %>%
  arrange(desc(total_sales))
```

#### Final Aggregated Digitec Dataset

The result of this cleaning and preprocessing workflow is the `canton_data_final` DataFrame. It contains 26 rows, one for each Swiss canton, with the key performance indicators required for further analysis (total number of sales, number of transactions, average value of orders, most ordered brand, most ordered category). This dataset is now ready to be merged with the BFS population data.

```{r}
#| label: display-final-table
#| echo: false

# Display the final, clean table
head(canton_data_final, n = 26)
```

### 4.2 BFS Datasets

```{r}

# Calculate number of transactions, average price, and median price for each bfsnr
digitec_grouped <- digitec_data_clean %>%
  group_by(bfsnr) %>%
  summarise(
    transactions = n(),
    avg_price = mean(price, na.rm = TRUE),
    median_price = median(price, na.rm = TRUE),
    .groups = "drop"
  )
```

## 5. Canton Dataset

### 5.1. Canton Level Dataset Merge

```{r}

# Merge on canton code
canton_merged_data <- left_join(canton_data_final, cleaned_canton_population, by = "canton")

# View merged dataset
canton_merged_data
```

### 5.2. Canton Level Post-merge Aggregation

To enable meaningful comparisons across cantons we add derived indicators. These indicators normalize raw counts into ratios and per-capita metrics, making them suitable for correlation analysis, segmentation, and strategic targeting.

The metrics are the following:

1.  Gender ratios
2.  Citizenship ratios
3.  Age structure
4.  Marital status

```{r}

 final_canton_data <- canton_merged_data %>% mutate( 
  
  # Economic indicators
  sales_per_capita = total_sales / total,
  transactions_per_capita = num_transactions / total,
  
  # Gender ratios
  share_male = male_population / total,

  # Citizenship ratios
  share_foreign = foreigner_citizenship / total,
  
  # Age structure
  share_age_0_19 = `age_from_0-19` / total,
  share_age_20_64 = `age_from_20-64` / total,
  share_age_65_plus = age_from_65_and_over / total,
  
  # Marital status
  share_single = single_status / total,
  share_married = married_status / total,
)
 
# Display final merged data with computed values (note: removal of original values.)
final_canton_data
```

## 6. Municipality Dataset

### 6.1 Municipality Level Dataset Merge

```{r}

# Merge age municipality dataset on bfsnr code
municipality_merged_data <- digitec_grouped %>%
  left_join(municipal_population_age_clean, by = "bfsnr")

# Merge gender municipality dataset on bfsnr code
municipality_merged_data <- municipality_merged_data %>%
  left_join(municipal_population_gender_clean, by = "bfsnr")

# View merged dataset
municipality_merged_data
```

### 6.2 Municipality Level Post-merge Aggregation

The metrics for the municipalities are the following:

1.  Gender ratios
2.  Citizenship ratios
3.  Age structure

```{r}

final_municipality_data <- municipality_merged_data %>%
  rename(
    male_population   = mann_all,
    female_population = frau_all,
    total_population  = total_all,
    swiss_citizenship    = total_swiss,
    foreigner_citizenship = total_foreigner
  ) %>%
  mutate(
    # choose one total population column to use everywhere
    total_pop = coalesce(total_population, total),

    # Gender ratios
    share_male   = male_population   / total_pop,

    # Citizenship ratios
    share_foreign = foreigner_citizenship / total_pop,
  
    # Age structure
    share_age_0_19    = `age_from_0-19`      / total_pop,
    share_age_20_64   = `age_from_20-64`     / total_pop,
    share_age_65_plus = age_from_65_and_over / total_pop
  )

# Display final merged data with computed values
final_municipality_data
```

## 7. Data visualization

We first want to fin out which categories have the widest distribution:

```{r}
#| label: category-price-spread
#| warning: false
#| message: false

library(dplyr)
library(forcats)
library(ggplot2)
library(scales)

# 1. Compute spread + transaction count per category
spread_stats <- digitec_data_clean |>
  group_by(category) |>
  summarise(
    n_obs        = n(),
    median_price = median(price, na.rm = TRUE),
    iqr_price    = IQR(price, na.rm = TRUE),
    sd_price     = sd(price, na.rm = TRUE),
    .groups = "drop"
  )

# 2. Rank categories by:
#    - widest IQR first
#    - within that, more transactions first
spread_stats_top <- spread_stats |>
  arrange(desc(iqr_price), desc(n_obs)) |>
  slice_head(n = 10)

top_spread_cats <- spread_stats_top$category

# 3. Plot boxplots for those top 10 categories
plot_spread <- digitec_data_clean |>
  filter(category %in% top_spread_cats) |>
  mutate(
    # order by IQR (descending) using the stats we just computed
    category = fct_reorder(
      category,
      price,
      .fun   = IQR,
      .desc  = TRUE
    )
  )

p_spread <- ggplot(plot_spread, aes(x = category, y = price)) +
  geom_boxplot(outlier.alpha = 0.15,
               fill = viridis::viridis(1, begin = 0.4, end = 0.6)) +
  coord_flip() +
  labs(
    title = "Top 10 Categories with Widest Price Distribution",
    subtitle = "Ranked by IQR (primary) and number of transactions (secondary)",
    x = "Category",
    y = "Price (CHF)"
  ) +
  theme_minimal()

p_spread



```

Lets focus on Notebook and PC and see if different cantons and municipalities prefer different prices!

### 7.1 Pre-visualization joins and processing

We first connect the cleaned Digitec transactions to the official municipality geometries.

```{r}
library(sf)
# Path 
muni <- st_read("swissBOUNDARIES/swissBOUNDARIES3D_1_5_TLM_HOHEITSGEBIET.shp")

# Inspect columns to find the BFS field
names(muni)

```

```{r}


library(sf)
library(dplyr)

canton_lookup <- tibble::tibble(
  KANTONSNUM = 1:26,
  canton = c(
    "ZH","BE","LU","UR","SZ","OW","NW","GL","ZG","FR",
    "SO","BS","BL","SH","AR","AI","SG","GR","AG","TG",
    "TI","VD","VS","NE","GE","JU"
  )
)

muni_canton <- muni %>%
  left_join(canton_lookup, by = "KANTONSNUM")

canton_sf <- muni_canton %>%
  group_by(canton) %>%
  summarise(geometry = st_union(geometry), .groups = "drop")

canton_sf <- st_cast(canton_sf, "MULTIPOLYGON")

```

```{r}
canton_map_data <- canton_sf %>%
  left_join(final_canton_data, by = "canton")

```

```{r}
#| label: canton-kpi-prep

library(dplyr)
library(stringr)
library(tidyr)

## 0.1 Cheap / Expensive segments

price_q <- quantile(data_final_swiss$price, probs = c(0.25, 0.75), na.rm = TRUE)
cheap_thr <- price_q[1]
exp_thr   <- price_q[2]

data_final_swiss <- data_final_swiss %>%
  mutate(
    price_segment = case_when(
      price <= cheap_thr ~ "Cheap",
      price >= exp_thr   ~ "Expensive",
      TRUE               ~ "Mid"
    )
  )

canton_segments <- data_final_swiss %>%
  count(canton, price_segment) %>%
  group_by(canton) %>%
  mutate(share = n / sum(n)) %>%
  ungroup() %>%
  pivot_wider(
    names_from  = price_segment,
    values_from = share,
    values_fill = 0
  ) %>%
  rename(
    cheap_share     = Cheap,
    mid_share       = Mid,
    expensive_share = Expensive
  )

final_canton_data <- final_canton_data %>%
  left_join(canton_segments, by = "canton")

data_final_swiss <- data_final_swiss %>%
  mutate(
    product_group = case_when(
      str_detect(category, "smartphone")                    ~ "Smartphone",
      str_detect(category, "gaming")                        ~ "Gaming",
      str_detect(category, "sportuhr-smartwatch")           ~ "Smartwatch",
      str_detect(category, "einbaukuehlschrank|foehn")      ~ "Home/Appliance",
      TRUE                                                  ~ "Other"
    )
  )

group_shares <- data_final_swiss %>%
  group_by(canton, product_group) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(canton) %>%
  mutate(share = n / sum(n)) %>%
  ungroup() %>%
  select(canton, product_group, share) %>%
  pivot_wider(
    names_from  = product_group,
    values_from = share,
    values_fill = 0
  ) %>%
  rename(
    smartphone_share = Smartphone,
    gaming_share     = Gaming,
    smartwatch_share = Smartwatch,
    home_share       = `Home/Appliance`
  )

final_canton_data <- final_canton_data %>%
  left_join(group_shares, by = "canton")

canton_map_data <- canton_sf %>%
  left_join(final_canton_data, by = "canton")
```

```{r}
library(dplyr)
library(stringr)

gaming_data <- data_final_swiss %>%
  filter(
    brandName == "Nintendo" |
    (brandName == "Ugreen" &
       str_detect(fullProductName, regex("switch|nintendo|joy[- ]?con", ignore_case = TRUE))) |
    str_detect(fullProductName, regex("playstation|ps4|ps5", ignore_case = TRUE)) |
    str_detect(fullProductName, regex("xbox", ignore_case = TRUE))
  ) %>%
  mutate(
    platform = case_when(
      brandName == "Nintendo" ~ "Nintendo",
      brandName == "Ugreen" &
        str_detect(fullProductName, regex("switch|nintendo|joy[- ]?con", ignore_case = TRUE)) ~ "Nintendo",
      str_detect(fullProductName, regex("playstation|ps4|ps5", ignore_case = TRUE)) ~ "PlayStation",
      str_detect(fullProductName, regex("xbox", ignore_case = TRUE)) ~ "Xbox",
      TRUE ~ "Other"
    )
  )
```

```{r}

gaming_canton <- gaming_data %>%
  group_by(canton) %>%
  summarise(
    gaming_sales        = sum(price, na.rm = TRUE),
    gaming_transactions = n(),
    avg_gaming_price    = mean(price, na.rm = TRUE),
    .groups = "drop"
  )

canton_map_data <- canton_map_data %>%
  filter(!is.na(canton)) %>%        
  select(-any_of(c("gaming_sales", 
                   "gaming_transactions", 
                   "avg_gaming_price", 
                   "tooltip_gaming",
                   "gaming_share_total"))) %>%  
  left_join(gaming_canton, by = "canton") %>%
  mutate(
    gaming_sales        = coalesce(gaming_sales, 0),
    gaming_transactions = coalesce(gaming_transactions, 0),
    avg_gaming_price    = coalesce(avg_gaming_price, 0),
    gaming_share_total  = if_else(total_sales > 0, 
                                  gaming_sales / total_sales, 
                                  0)
  ) %>%
  mutate(
    tooltip_gaming = paste0(
      "<b>Canton: </b>", canton,
      "<br><b>Gaming sales:</b> ", scales::comma(round(gaming_sales, 0)), " CHF",
      "<br><b>Transactions:</b> ", gaming_transactions,
      "<br><b>Avg basket:</b> ", round(avg_gaming_price, 1), " CHF",
      "<br><b>Gaming share:</b> ",
      scales::percent(gaming_share_total, accuracy = 0.1)
    )
  )

```

### 7.2 Gaming sales across Switzerland

To understand regional differences in gaming demand, we first aggregated Nintendo, PlayStation, Xbox and Ugreen (Switch-related) transactions to the canton level. We then merged these data with demographic and geographic information from the Federal Statistical Office.

The interactive map below visualizes total gaming sales by canton. Hovering over a canton displays key metrics, including total gaming sales, number of transactions, average gaming basket size, and the canton’s gaming share.

```{r}
library(ggplot2)
library(plotly)
library(viridisLite)

p_gaming_sales <- ggplot(canton_map_data) +
  geom_sf(
    aes(
      fill = gaming_sales,
      text = tooltip_gaming
    ),
    color = "grey20",
    linewidth = 0.3
  ) +
  scale_fill_viridis_c(
    option = "C",
    name = "Gaming Sales [CHF]",
    labels = scales::label_comma()
  ) +
  labs(
    title = "Gaming Sales by Canton",
    subtitle = "Nintendo, PlayStation, Xbox & Ugreen (Switch)"
  ) +
  theme_minimal()

ggplotly(p_gaming_sales, tooltip = "text")

```

Gaming demand in Switzerland is heavily concentrated in urban cantons. Zurich is the single most important market for Digitec’s gaming category, followed by Bern, Aargau, Vaud, St. Gallen, and Geneva. These cantons not only drive the majority of gaming sales but also include the highest-density gamer populations based on gaming share percentage.

Cantons with zero or low gaming activity indicate rural areas or small populations where outreach will not significantly impact total sales. So, if Digitec wants to maximize return on advertising spend, focusing on the six large cantons gives access to the largest clusters of active gamers.

```{r}
top_cantons <- canton_map_data %>%
  distinct(canton, .keep_all = TRUE) %>%
  arrange(desc(gaming_sales)) %>%
  slice_head(n = 6) %>%
  pull(canton)

top_cantons

```

```{r}
gaming_top <- gaming_data %>%
  filter(canton %in% top_cantons)

platform_canton <- gaming_top %>%
  group_by(canton, platform) %>%
  summarise(
    sales = sum(price, na.rm = TRUE),
    transactions = n(),
    avg_price = mean(price, na.rm = TRUE),
    .groups = "drop"
  )

```

We next examined which gaming platforms dominate in the six strongest cantons.

```{r}
library(ggplot2)
library(scales)

ggplot(platform_canton, aes(x = canton, y = sales, fill = platform)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = comma) +
  scale_fill_viridis_d(option = "C") +
  labs(
    title = "Platform Breakdown in Top Gaming Cantons",
    x = "Canton",
    y = "Gaming Sales [CHF]",
    fill = "Platform"
  ) +
  theme_minimal(base_size = 14)
```

In **five out of six** of these top cantons (AG, BE, GE, SG, ZH), PlayStation generates the highest sales. Xbox is consistently the smallest of the three but still meaningful everywhere, especially in Zurich and Aargau. **Vaud (VD) is the exception where Nintendo dominates**.

In short, Digitec should push PlayStation heavily in Zürich and the German-speaking top cantons where it dominates, while tailoring Vaud with more Switch-focused offers.

We found it very interesting that Vaud (VD) is the only top canton where Nintendo beats PlayStation in sales. A Nintendo-specific regional campaign will perform better here than a broad PlayStation related campaigns. This makes it Digitec’s strongest opportunity for a targeted Switch-focused campaign. By breaking down Nintendo sales inside Vaud’s municipalities, we can identify exactly where the most engaged family and casual gamers live.

So the *right question* now is:

**Why is Vaud different? Who actually buys all this Nintendo equipment?**

To understand why Nintendo is so strong in Vaud, we analyzed how gaming purchasing relates to the canton’s age structure.

```{r}
vaud_demo <- canton_map_data %>%
  filter(canton == "VD") %>%
  slice(1) %>%   # <-- IMPORTANT
  mutate(
    share_age_0_19 = `age_from_0-19` / total,
    share_age_20_64 = `age_from_20-64` / total,
    share_age_65_plus = age_from_65_and_over / total
  )
vaud_nintendo <- gaming_data %>%
  filter(canton == "VD", platform == "Nintendo") %>%
  select(fullProductName, price)

total_nintendo_sales_vaud <- sum(vaud_nintendo$price, na.rm = TRUE)

vaud_nintendo_age <- tibble(
  age_group = c("0_19", "20_64", "65_plus"),
  sales = c(
    total_nintendo_sales_vaud * vaud_demo$share_age_0_19,
    total_nintendo_sales_vaud * vaud_demo$share_age_20_64,
    total_nintendo_sales_vaud * vaud_demo$share_age_65_plus
  )
)


```

```{r}
total_nintendo_sales_vaud <- sum(vaud_nintendo$price, na.rm = TRUE)

vaud_nintendo_age <- tibble(
  age_group = c("0_19", "20_64", "65_plus"),
  sales = c(
    total_nintendo_sales_vaud * vaud_demo$share_age_0_19,
    total_nintendo_sales_vaud * vaud_demo$share_age_20_64,
    total_nintendo_sales_vaud * vaud_demo$share_age_65_plus
  )
)

ggplot(vaud_nintendo_age,
       aes(x = age_group, y = sales)) +
  geom_col(fill = "#FF4B4B") +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Nintendo Sales in Vaud by Age Segment (Canton Level)",
    x = "Age Group",
    y = "Nintendo Sales [CHF]"
  ) +
  theme_minimal(base_size = 14)

```

The 20–64 age segment generates the majority of Nintendo spending. Youth (0–19) contribute strongly, while seniors play only a minor role.

This means Nintendo’s strength in Vaud comes from **adult players buying premium products** or from parents purchasing consoles for children. But what do they buy? Do they buy products for themselves or their children?

```{r}
vaud_demo <- canton_map_data %>%
  filter(canton == "VD") %>%
  slice(1) %>%  
  mutate(
    share_age_20_64 = `age_from_20-64` / total
  )

vaud_nintendo <- gaming_data %>%
  filter(canton == "VD", platform == "Nintendo") %>%
  select(fullProductName, price)

vaud_nintendo_weighted <- vaud_nintendo %>%
  mutate(
    weighted_sales = price * vaud_demo$share_age_20_64
  )

nintendo_product_summary <- vaud_nintendo_weighted %>%
  group_by(fullProductName) %>%
  summarise(
    sales_20_64 = sum(weighted_sales, na.rm = TRUE),
    transactions = n(),
    .groups = "drop"
  ) %>%
  arrange(desc(sales_20_64)) %>%
  slice_head(n = 10)  


```

```{r}
library(ggplot2)
library(scales)
library(stringr)

#for visualization purposes
nintendo_product_summary$fullProductName <- str_wrap(nintendo_product_summary$fullProductName, width = 30)

ggplot(nintendo_product_summary,
       aes(x = reorder(fullProductName, sales_20_64),
           y = sales_20_64)) +
  geom_col(fill = "#E60012") +  # Nintendo red
  coord_flip() +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Top Nintendo Products Bought by Adults (20–64) in Vaud",
    x = "Nintendo Product",
    y = "Weighted Sales [CHF]"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.text.y = element_text(size = 9), 
    plot.margin = margin(10, 20, 10, 20)     
  )


```

The top items in the chart are almost all **premium editions of switch consoles or games**. This shows that adults in Vaud are making **big Nintendo purchases**, not just small add-ons.

Vaud’s adult population buys premium Nintendo consoles and major franchises like Zelda. Zelda is a franchise primarily popular among adults, so we can say that a large portion of Nintendo customers in Vaud are adult gamers, not only parents shopping for kids as some people might think.

This means Nintendo advertising in Vaud should not focus only on kids but it must mostly target adult gamers who choose Switch for themselves.

#### Summary

Zurich, Bern, Aargau, St. Gallen, and Geneva all show a clear preference for PlayStation. These cantons are also the largest contributors to gaming sales overall. So, Digitec should push **PS5 bundles, controllers, PS Plus subscriptions, and headline releases** heavily in these markets.

Vaud is the only major canton where Nintendo dominates PlayStation. Adults (20–64) drive most of these sales, especially premium Switch models and Zelda titles. Campaigns here should focus on **adult Nintendo players**, premium Switch upgrades and promotions on favourite fanchises like Zelda.
