---
title: "Digitec Analysis"
format:
  html:
    df-print: paged
    toc: true
    code-tools: true

editor: visual
author: Fereshteh, Anna, Sen, Colin Marti (pls add your full name)
---

## 1. Business Plan Analysis

Digitec Galaxus is Switzerland’s largest online retailer for consumer electronics, operating exclusively via e-commerce. Despite nationwide reach, customer behavior likely differs across cantons due to income, urbanization, age, citizenship, and household structures.

This project explores how these demographic and regional factors influence online purchases using: - \~163,000 Digitec transaction records, and - Canton-level demographic data from the Swiss Federal Statistical Office (BFS).

The findings will guide a data-driven marketing strategy to personalize digital campaigns and increase online sales efficiency.

### 1.1. Business Problem

> **How can regional demographics be used to improve Digitec’s online marketing strategy and boost conversions across Switzerland?**

Digitec currently uses uniform national marketing, but demographic diversity suggests that: - Urban cantons might favor high-end tech, - Rural areas could prefer affordable products, - Younger, single populations may show distinct category interests.

By linking demographic patterns with online purchasing data, Digitec can **target audiences more intelligently** and tailor its digital campaigns for maximum impact.

### 1.2. Data Overview

| Source | Description | Key Variables |
|:-----------------------|:-----------------------|:-----------------------|
| **Digitec Transactions (`DigitecLive_Final.csv`)** | \~163k web-scraped online transactions from Digitec’s live purchase feed | `brandName`, `infos.Category`, `salesPrice.amountIncl`, `dateTime`, `mun_canton` |
| **BFS Population Data (`population_2023.xlsx`)** | Canton-level demographics (2023) | `Total`, `Swiss`, `Foreigner`, `Urban`, `Intermediate`, `Rural`, `Age_0_19`, `Age_20_64`, `Age_65plus`, `Single`, `Married`, `Widowed`, `Divorced` |
| **BFS Population by Age (`mns-population-by-age-2023.xlsx`)** | Detailed population structure by age and region | `Region`, `Total`, `Age_0_4`, `Age_5_9`, …, `Age_80plus` |
| **BFS Population by Gender (`mns-population-by-gender-2023.xlsx`)** | Gender composition by canton and municipality (2023) | `Region`, `Total_All`, `Mann_All`, `Frau_All`, `Swiss`, `Foreigner` |
| **Swiss Municipal Boundaries (`swissBOUNDARIES3D_1_5_TLM_HOHEITSGEBIET.shp`)** | Official Swiss municipality polygons used to map Digitec transactions geographically and join by BFS municipality number | `BFS_NUMMER` (municipality ID), `NAME` (municipality name), `KANTONSNUM` (canton ID), `geometry` (MULTIPOLYGON) |

These additional datasets allow the inclusion of **more granular demographic information** — enabling a refined analysis of age- and gender-related purchasing behavior across regions.

### 1.3. Unit of Analysis

The analysis is now conducted at the **municipality level** ($n \approx 2,200$), using the official **BFS municipality number (`mun_bfsnr`)** as the join key between datasets.

Each municipality represents one observation containing: - $sales\_per\_capita = \frac{total\_sales}{population}$\
- Average order value (CHF)\
- Age-group shares (e.g., % under 20, % over 65)\
- Gender composition (% male, % female)\
- Citizenship composition (% Swiss, % foreign)

This finer-grained approach allows for **more localized insights**, capturing demographic and purchasing differences that are hidden when aggregating by canton.

------------------------------------------------------------------------

### 1.4. Methodology

#### 1.4.1. Analytical Steps

1.  **Aggregate** Digitec transaction data by municipality (`mun_bfsnr`)\
    → Compute total sales, average price, and transaction count for each municipality.\
2.  **Merge** the municipality-level BFS datasets on population, gender, and age using the `mun_bfsnr` key.\
3.  **Compute derived demographic indicators:**
    -   Gender ratio: $\frac{Male}{Female}$\
    -   Share of foreign residents: $\frac{Foreigner}{Total}$\
    -   Age structure: $\frac{Population_{0-19}}{Total}$, $\frac{Population_{20-64}}{Total}$, $\frac{Population_{65+}}{Total}$\
    -   Sales per capita and average order value\
4.  **Explore correlations** between online sales indicators and demographics:
    -   Are younger or more international municipalities buying more expensive products?\
    -   Do gender or age structures explain variations in spending per capita?\
5.  **Visualize** geographic and demographic patterns using:
    -   Choropleth maps by municipality (sales per capita, average price, etc.)\
    -   Scatterplots linking demographics to sales variables\
    -   Heatmaps of demographic vs. economic indicators\
6.  **Identify four key insights** to inform Digitec’s new data-driven marketing campaign.

#### 1.4.2. Research Questions

| **Demographic Dimension** | **Research Question** | **Usage Example (What Digitec Could Do)** |
|------------------------|------------------------|------------------------|
| **Population Size** | Do larger cantons generate more online purchases? | Focus major campaigns and product launches in high-activity cantons, while using localized ads to increase engagement in smaller regions. |
| **Gender Composition** | Does the male/female balance influence which product categories are bought? | Promote gaming content in female-dominant cantons and highlight smartphone or smartwatch deals in male-dominant regions. |
| **Citizenship** | Do cantons with more foreign residents show different tech-buying patterns? | Use multilingual ads (DE/FR/EN) and highlight global brands such as Apple or Samsung in foreign-dense cantons. |
| **Urbanization** | Do urban regions buy more high-end devices compared to rural regions? | Promote flagships and new releases in urban cantons; emphasize budget or refurbished products in rural areas. |
| **Marital Status** | Do single vs. married populations differ in what they purchase? | Focus gaming offers in single-dominant cantons; promote smart-home and smartwatch bundles in married-dominant regions. |

### 1.5. Strategic Impact

By merging **online sales data** with **BFS demographic indicators**, Digitec can: - Move from national-level campaigns to **regionally adaptive marketing**.\
- Align ad language, pricing, and products with **local audience characteristics**.\
- Strengthen relevance and **optimize ad spend efficiency**.

## 2. Setup and Data Loading

This first code chunk sets up the R environment by loading all necessary libraries for the analysis and loading the four primary datasets. The chunk is configured with `include: false` to hide it from the final rendered report for a cleaner presentation.

```{r}

#| label: setup
#| include: false

# Load all required libraries for the project
library(tidyverse) # For data manipulation (dplyr) and visualization (ggplot2)
library(readxl)    # To read Excel files
library(lubridate) # To work with date-time values
library(stringr)   # To work with regex
library(dlookr)
library(ggrepel)
library(ggcorrplot)
library(viridis)
library(dplyr)
library(plotly)
library(sf)
library(scales)    

# --- GLOBAL DESIGN DEFINITION ---
# Save the original theme_minimal
orig_theme_minimal <- ggplot2::theme_minimal

# Override with custom settings
theme_minimal <- function(base_size = 11, base_family = "", ...) {
  orig_theme_minimal(base_size = base_size, base_family = base_family, ...) +
  theme(
    # 1. Axis text (values on ticks)
    axis.text = element_text(size = 18, color = "black"),
    
    # 2. Axis titles 
    axis.title = element_text(size = 20, color = "black", margin = margin(t=15, r=15)),
    
    # 3. Main Plot Title
    plot.title = element_text(size = 26, color = "black", margin = margin(b=15)),
    
    # 4. Plot Subtitle
    plot.subtitle = element_text(
      size = 20, 
      color = "black", 
      margin = margin(b=20)
    ),
    
    # 5. Caption 
    plot.caption = element_text(size = 14, color = "grey30", hjust = 1),
    
    # 6. Legends
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18)
  )
}

# Load digitect dataset
load("digitec.RData") # Loads the 'data' object

# Save to descriptive variable
digitec_data <- get("data")

## Load BFS data sets
canton_population <- read_excel("population_2023.xlsx")
municipal_population_gender <- read_excel("mns-population-by-gender-2023.xlsx")
municipal_population_age <- read_excel("mns-population-by-age-2023.xlsx")

```

## 3. Data Cleaning and Preprocessing

This section details the systematic process of cleaning and preparing the raw data of the required data sets. The goal is to produce a reliable, aggregated dataset at the municipal and cantonal level. The municipal data will be a combination of three datasets: "mns-population-by-age", "mns-population-by-gender", and the digitec dataset. The cantonal dataset will be a merging of the BFS demographic data and the digitec dataset.

### 3.1. Digitec Dataset

An initial inspection of the raw Digitec data revealed several necessary cleaning steps:

-   **Correcting Data Types:** The `dateTime` column was stored as a character string and needed to be converted into a proper date-time format for time-based analysis.
-   **Removing Irrelevant Columns:** Many columns were redundant (e.g., price with and without VAT), contained technical metadata, or were not relevant to the analysis (e.g., image URLs).
-   **Ensuring Data Quality:** Transactions without a price or a clear, unique geographic match (`mun_nhits != 1`) were removed to ensure the reliability of the analysis.
-   **Feature Engineering:** More useful features were extracted from the raw data, such as the day of the week from the `dateTime` column or the clean product category name from a URL string.

#### 3.1.1. Data Cleaning & Engineering

First we check the dataset columns that we think might require some filtering or deletion.

```{r}
candidates <- c(
  "cityName",
  "salesPrice.currency",
  "infos.Price",
  "infos.Currency",
  "salesPrice",
  "mun_nhits"
)

existing <- intersect(candidates, names(digitec_data))

check_var <- function(df, var) {
  x <- df[[var]]
  tibble(
    variable      = var,
    exists        = TRUE,
    n             = length(x),
    n_non_na      = sum(!is.na(x)),
    prop_na       = mean(is.na(x)),
    n_unique_non_na = dplyr::n_distinct(x, na.rm = TRUE),
    example_values  = paste(utils::head(unique(x), 3), collapse = " | ")
  )
}

profile <- map_dfr(existing, ~check_var(digitec_data, .x))
profile
```

-   `salesPrice.currency`, `infos.Currency` → `n_unique_non_na == 1` (“CHF”) we decided to remove this column.
-   `salesPrice` → `n_non_na == 0` empty column so we decided to delete it.
-   `mun_nhits` → mostly/all 1

```{r}
if ("mun_nhits" %in% names(digitec_data)) {
  cat("mun_nhits frequency (percent):\n")
  perc <- prop.table(table(digitec_data$mun_nhits, useNA = "ifany")) * 100
  print(round(perc, 2))
}
```

-   **`1`** → 153 778 or 94.35% of the rows matched exactly one municipality.
-   **`0`** → 9 194 or 5.64% of the rows had *no municipality match*.
-   **`2`** → 21 or 0.01% of the rows matched *two possible municipalities*.

We would drop the rows where `mun_nhits == 0` or `mun_nhits == 2`, then remove the `mun_nhits` column. Since we already have 95%.

-   `cityName` → unique city name inserted by customer, since we have the city and municipality takes from SWISS POST Api we decided to drop this column.
-   `infos.Price` → many NAs - 63% of the data is missing so the column is invalid.

```{r}
if (all(c("infos.Price", "salesPrice.amountIncl") %in% names(digitec_data))) {
  same_info <- digitec_data %>%
    filter(!is.na(infos.Price) & !is.na(salesPrice.amountIncl)) %>%
    summarise(
      n = n(),
      max_abs_diff = max(abs(infos.Price - salesPrice.amountIncl), na.rm = TRUE)
    )
  same_info
}
```

60 577 rows had *both* `infos.Price` and `salesPrice.amountIncl` . The value `max_abs_diff = 1021.9` means that in some rows, the difference between the two price fields is CHF 1 021.90 so prices differ a lot. We decided that this column is inconsistent.

The following steps were implemented in a single, coherent `dplyr` pipeline for efficiency and reproducibility. The result is a cleaned DataFrame named `data_clean`.

```{r}
#| label: data-cleaning-pipeline

library(dplyr)
library(tidyr)
library(lubridate)
library(forcats)

digitec_data_clean <- digitec_data %>%
  # 1) Select and rename relevant columns for better readability
  rename(
    price        = salesPrice.amountIncl,
    category_url = infos.Category,
    sales_rank_raw = infos.Sales_Rank,
    canton       = mun_canton,
    bfsnr        = mun_bfsnr
  ) %>%
  select(
    price, dateTime, brandName, fullProductName,
    socialShoppingTransactionTypeId,
    category_url, sales_rank_raw, canton, mun_nhits, bfsnr
  ) %>%
  
  # 2) Filter rows to ensure data quality
  filter(
    !is.na(price),                  # keep only transactions with a price
    !is.na(canton),                 # keep only transactions with canton info
    mun_nhits == 1,                 # keep only unique geographic matches
    socialShoppingTransactionTypeId %in% c(3, 5),  # only bought/shipped
    grepl("von", sales_rank_raw)    # keep only rows where rank contains "von"
  ) %>%
  
  # 3) Remove irrelevant or redundant columns
  select(-any_of(c(
    "cityName",
    "salesPrice.currency",
    "infos.Price",
    "infos.Currency",
    "salesPrice"
  ))) %>%
  
  # 4) Create new, useful columns (feature engineering)
  mutate(
    # timestamps
    date     = as_date(dateTime),
    dateTime = ymd_hms(dateTime),
    hour     = hour(dateTime),
    weekday  = wday(dateTime, label = TRUE, week_start = 1),
    # product category from URL
    category = basename(category_url)
  ) %>%
  
  # 5) Parse sales rank: "12 von 345" -> numeric position / total
  separate(
    sales_rank_raw,
    into = c("rank_position", "rank_total"),
    sep  = " von "
  ) %>%
  mutate(
    rank_position = as.numeric(rank_position),
    rank_total    = as.numeric(rank_total),
    rank_relative = rank_position / rank_total,
    rank_score    = 1 - rank_relative
  ) %>%
  
  # 6) Recode transaction type and convert key vars to factors
  mutate(
    transaction_type = case_when(
      socialShoppingTransactionTypeId == 3 ~ "bought",
      socialShoppingTransactionTypeId == 5 ~ "shipped"
    ),
    transaction_type = fct_relevel(transaction_type, "bought", "shipped"),
    category  = as.factor(category),
    brandName = as.factor(brandName)
  ) %>%
  select(-socialShoppingTransactionTypeId, -mun_nhits) %>%  # no longer needed
  
  # 7) Drop duplicates
  distinct() %>%
  
  # 8) Handle Inf and winsorize rank metrics in one go
  mutate(
    rank_score    = ifelse(is.infinite(rank_score),    NA_real_, rank_score),
    rank_relative = ifelse(is.infinite(rank_relative), NA_real_, rank_relative)
  ) %>%
  { 
    qs_score <- quantile(.$rank_score,    probs = c(0.025, 0.975), na.rm = TRUE)
    qs_rel   <- quantile(.$rank_relative, probs = c(0.025, 0.975), na.rm = TRUE)
    
    mutate(.,
      rank_score    = pmin(pmax(rank_score,    qs_score[1]), qs_score[2]),
      rank_relative = pmin(pmax(rank_relative, qs_rel[1]),   qs_rel[2])
    )
  } %>%
  filter(!is.na(rank_score), !is.na(rank_relative))
```

```{r}
overview(digitec_data_clean)
summary(digitec_data_clean)

```




### 3.2. BFS Datasets

The data within the three BFS is mostly clean. There are no apparent missing values or errors, just some unnecessary rows. More importantly, the datasets need to be compatible for the upcoming merge with the Digitec data.\
\
The issues that we identified were as follows:

-   **Column names:** The column names of all three datasets need to be non-capitalized and follow the snake_case format in order to align with the digitec dataset.

-   **Unwanted rows:** Remove the unwanted canton and bezirk rows with the municipality datasets and rename the column to muncipality.

-   **Canton codes:** The canton names within the municipality BFS dataset are not in the standard code format.

-   **Municipality number:** The age and gender datasets have an improper municipality notation. Later on, we'll have to merge these datasets with the digitec dataset according to the municipality number, so it's imperative that we make sure they're in the same format.

-   **Ages:** Currently ages within the age datasetare not grouped in the same way as BFS' canton dataset.

#### 3.2.1 Data Cleaning & Engineering

**Municipality Level**

Firstly, we will tackle the column names; we do this simply by running each dataset through this snake_case function:

```{r}

#function which converts input to snake_case
clean_names <- function(x) {
  x %>%
    str_to_lower() %>%
    
    # Replace spaces with underscores
    str_replace_all(" ", "_")             
}

# Rename each dataset
colnames(canton_population) <- clean_names(colnames(canton_population))
colnames(municipal_population_age) <- clean_names(colnames(municipal_population_age))
colnames(municipal_population_gender) <- clean_names(colnames(municipal_population_gender))
```

Then within the age and gender datasets we can remove the unnecessary rows, rename the column and extract the bfsnr from the municipalities.

```{r}
#| label: row-removal

# Rename first column to bfsnr like in the digitec dataset
colnames(municipal_population_age)[1] <- "bfsnr"
colnames(municipal_population_gender)[1] <- "bfsnr"

# Clean municipality age data
municipal_population_age_clean <- municipal_population_age %>%
  filter(str_detect(bfsnr, "^\\.+")) %>%  # Match rows starting with one or more dots
  mutate(
    bfsnr = str_extract(bfsnr, "\\d+"),  # Extract numeric code
    bfsnr = as.integer(bfsnr)
  )

# Clean municipality gender data
municipal_population_gender_clean <- municipal_population_gender %>%
  filter(str_detect(bfsnr, "^\\.+")) %>%
  mutate(
    bfsnr = str_extract(bfsnr, "\\d+"),
    bfsnr = as.integer(bfsnr)
  )

municipal_population_age_clean
municipal_population_gender_clean

```

Now we group the ages from 0-19, 20-64, and 65-100+ and our age and population datasets are ready for merging.

```{r}

# Collapse into age bands
municipal_population_age_clean <- municipal_population_age_clean %>%
  rowwise() %>%
  mutate(
    `age_from_0-19` = sum(c_across(`0`:`19`), na.rm = TRUE),
    `age_from_20-64` = sum(c_across(`20`:`64`), na.rm = TRUE),
    age_from_65_and_over = sum(c_across(`65`:`99`), `100_und_mehr`, na.rm = TRUE)
  ) %>%
  ungroup() %>% select(bfsnr, total, `age_from_0-19`, `age_from_20-64`, age_from_65_and_over)

municipal_population_age_clean
```

**Canton Level**

For the canton level we need to map the canton names found in the canton BFS dataset to their official abbreviations used in the transaction dataset.

```{r}
#| label: canton-code-mapping


# Define canton name-to-code mapping
canton_codes <- list(
  "Zurich" = "ZH",
  "Bern" = "BE",
  "Lucerne" = "LU",
  "Uri" = "UR",
  "Schwyz" = "SZ",
  "Obwalden" = "OW",
  "Nidwalden" = "NW",
  "Glarus" = "GL",
  "Zug" = "ZG",
  "Fribourg" = "FR",
  "Solothurn" = "SO",
  "Basel-Stadt" = "BS",
  "Basel-Landschaft" = "BL",
  "Schaffhausen" = "SH",
  "Appenzell A. Rh." = "AR",
  "Appenzell I. Rh." = "AI",
  "St. Gallen" = "SG",
  "Graubünden" = "GR",
  "Aargau" = "AG",
  "Thurgau" = "TG",
  "Ticino" = "TI",
  "Vaud" = "VD",
  "Valais" = "VS",
  "Neuchâtel" = "NE",
  "Geneva" = "GE",
  "Jura" = "JU"
)

# Convert to named vector for mapping
code_map <- setNames(unlist(canton_codes), names(canton_codes))

# Replace canton names with codes and remove non-cantons
cleaned_canton_population <- canton_population %>%
  filter(canton %in% names(canton_codes)) %>%
  mutate(canton = code_map[canton])


# Display the processed excel data
cleaned_canton_population

```

## 4. Validation and Final Aggregation

### 4.1 Digitec Dataset

An initial aggregation of the cleaned data surprisingly resulted in 29 groups instead of the 26 expected Swiss cantons. An investigation of the `canton` column revealed that the dataset also contained transactions from neighboring countries, which needed to be excluded for this analysis.

```{r}
#| label: validation-step

# A list of valid Swiss canton abbreviations
valid_swiss_cantons <- c(
  "AG", "AI", "AR", "BE", "BL", "BS", "FR", "GE", "GL", "GR", 
  "JU", "LU", "NE", "NW", "OW", "SG", "SH", "SO", "SZ", "TG", 
  "TI", "UR", "VD", "VS", "ZG", "ZH"
)

# Identify which unique codes in our data are NOT in the valid list
non_swiss_codes <- setdiff(unique(digitec_data_clean$canton), valid_swiss_cantons)

print("Found the following non-Swiss location codes:")
non_swiss_codes
```

Based on this finding, the dataset was filtered to include only transactions from the 26 Swiss cantons. Afterward, the final aggregation was performed to create the analysis-ready dataset.

```{r}
#| label: final-aggregation

# Filter the dataset to retain only transactions from Swiss cantons
data_final_swiss <- digitec_data_clean %>%
  filter(canton %in% valid_swiss_cantons)

# Perform the final aggregation at the cantonal level
canton_data_final <- data_final_swiss %>%
  group_by(canton) %>%
  summarise(
    total_sales = sum(price, na.rm = TRUE),
    num_transactions = n(),
    avg_order_value = mean(price, na.rm = TRUE),
    top_brand = names(which.max(table(brandName))),
    top_category = names(which.max(table(category)))
  ) %>%
  arrange(desc(total_sales))
```

#### Final Aggregated Digitec Dataset

The result of this cleaning and preprocessing workflow is the `canton_data_final` DataFrame. It contains 26 rows, one for each Swiss canton, with the key performance indicators required for further analysis (total number of sales, number of transactions, average value of orders, most ordered brand, most ordered category). This dataset is now ready to be merged with the BFS population data.

```{r}
#| label: display-final-table
#| echo: false

# Display the final, clean table
head(canton_data_final, n = 26)
```

### 4.2 BFS Datasets

```{r}

# Calculate number of transactions, average price, and median price for each bfsnr
digitec_grouped <- digitec_data_clean %>%
  group_by(bfsnr) %>%
  summarise(
    transactions = n(),
    avg_price = mean(price, na.rm = TRUE),
    median_price = median(price, na.rm = TRUE),
    .groups = "drop"
  )
```

## 5. Canton Dataset

### 5.1. Canton Level Dataset Merge

```{r}

# Merge on canton code
canton_merged_data <- left_join(canton_data_final, cleaned_canton_population, by = "canton")

# View merged dataset
canton_merged_data
```

### 5.2. Canton Level Post-merge Aggregation

To enable meaningful comparisons across cantons we add derived indicators. These indicators normalize raw counts into ratios and per-capita metrics, making them suitable for correlation analysis, segmentation, and strategic targeting.

The metrics are the following:

1.  Gender ratios
2.  Citizenship ratios
3.  Age structure
4.  Marital status

```{r}

 final_canton_data <- canton_merged_data %>% mutate( 
  
  # Economic indicators
  sales_per_capita = total_sales / total,
  transactions_per_capita = num_transactions / total,
  
  # Gender ratios
  share_male = male_population / total,

  # Citizenship ratios
  share_foreign = foreigner_citizenship / total,
  
  # Age structure
  share_age_0_19 = `age_from_0-19` / total,
  share_age_20_64 = `age_from_20-64` / total,
  share_age_65_plus = age_from_65_and_over / total,
  
  # Marital status
  share_single = single_status / total,
  share_married = married_status / total,
)
 
# Display final merged data with computed values (note: removal of original values.)
final_canton_data
```

## 6. Municipality Dataset

### 6.1 Municipality Level Dataset Merge

```{r}

# Merge age municipality dataset on bfsnr code
municipality_merged_data <- digitec_grouped %>%
  left_join(municipal_population_age_clean, by = "bfsnr")

# Merge gender municipality dataset on bfsnr code
municipality_merged_data <- municipality_merged_data %>%
  left_join(municipal_population_gender_clean, by = "bfsnr")

# View merged dataset
municipality_merged_data
```

### 6.2 Municipality Level Post-merge Aggregation

The metrics for the municipalities are the following:

1.  Gender ratios
2.  Citizenship ratios
3.  Age structure

```{r}

final_municipality_data <- municipality_merged_data %>%
  rename(
    male_population   = mann_all,
    female_population = frau_all,
    total_population  = total_all,
    swiss_citizenship    = total_swiss,
    foreigner_citizenship = total_foreigner
  ) %>%
  mutate(
    # choose one total population column to use everywhere
    total_pop = coalesce(total_population, total),

    # Gender ratios
    share_male   = male_population   / total_pop,

    # Citizenship ratios
    share_foreign = foreigner_citizenship / total_pop,
  
    # Age structure
    share_age_0_19    = `age_from_0-19`      / total_pop,
    share_age_20_64   = `age_from_20-64`     / total_pop,
    share_age_65_plus = age_from_65_and_over / total_pop
  )

# Display final merged data with computed values
final_municipality_data
```

#### 7. Data visualization

We first want to fin out which categories have the widest distribution:
```{r}
#| label: category-price-spread
#| warning: false
#| message: false

library(dplyr)
library(forcats)
library(ggplot2)
library(scales)

# 1. Compute spread + transaction count per category
spread_stats <- digitec_data_clean |>
  group_by(category) |>
  summarise(
    n_obs        = n(),
    median_price = median(price, na.rm = TRUE),
    iqr_price    = IQR(price, na.rm = TRUE),
    sd_price     = sd(price, na.rm = TRUE),
    .groups = "drop"
  )

# 2. Rank categories by:
#    - widest IQR first
#    - within that, more transactions first
spread_stats_top <- spread_stats |>
  arrange(desc(iqr_price), desc(n_obs)) |>
  slice_head(n = 10)

top_spread_cats <- spread_stats_top$category

# 3. Plot boxplots for those top 10 categories
plot_spread <- digitec_data_clean |>
  filter(category %in% top_spread_cats) |>
  mutate(
    # order by IQR (descending) using the stats we just computed
    category = fct_reorder(
      category,
      price,
      .fun   = IQR,
      .desc  = TRUE
    )
  )

p_spread <- ggplot(plot_spread, aes(x = category, y = price)) +
  geom_boxplot(outlier.alpha = 0.15,
               fill = viridis::viridis(1, begin = 0.4, end = 0.6)) +
  coord_flip() +
  labs(
    title = "Top 10 Categories with Widest Price Distribution",
    subtitle = "Ranked by IQR (primary) and number of transactions (secondary)",
    x = "Category",
    y = "Price (CHF)"
  ) +
  theme_minimal()

p_spread



```

Lets focus on Notebook and PC and see if different cantons and municipalities prefer different prices!




## 7.1 Visualization Cantons

We first connect the cleaned Digitec transactions to the official municipality geometries.

```{r}
library(sf)
# Path 
muni <- st_read("swissBOUNDARIES/swissBOUNDARIES3D_1_5_TLM_HOHEITSGEBIET.shp")

# Inspect columns to find the BFS field
names(muni)

```

### 7.1.1  pre-visualization joins and processing

```{r}


library(sf)
library(dplyr)

canton_lookup <- tibble::tibble(
  KANTONSNUM = 1:26,
  canton = c(
    "ZH","BE","LU","UR","SZ","OW","NW","GL","ZG","FR",
    "SO","BS","BL","SH","AR","AI","SG","GR","AG","TG",
    "TI","VD","VS","NE","GE","JU"
  )
)

muni_canton <- muni %>%
  left_join(canton_lookup, by = "KANTONSNUM")

canton_sf <- muni_canton %>%
  group_by(canton) %>%
  summarise(geometry = st_union(geometry), .groups = "drop")

canton_sf <- st_cast(canton_sf, "MULTIPOLYGON")

```

```{r}
canton_map_data <- canton_sf %>%
  left_join(final_canton_data, by = "canton")

```

```{r}
#| label: canton-kpi-prep

library(dplyr)
library(stringr)
library(tidyr)

## 0.1 Cheap / Expensive segments

price_q <- quantile(data_final_swiss$price, probs = c(0.25, 0.75), na.rm = TRUE)
cheap_thr <- price_q[1]
exp_thr   <- price_q[2]

data_final_swiss <- data_final_swiss %>%
  mutate(
    price_segment = case_when(
      price <= cheap_thr ~ "Cheap",
      price >= exp_thr   ~ "Expensive",
      TRUE               ~ "Mid"
    )
  )

canton_segments <- data_final_swiss %>%
  count(canton, price_segment) %>%
  group_by(canton) %>%
  mutate(share = n / sum(n)) %>%
  ungroup() %>%
  pivot_wider(
    names_from  = price_segment,
    values_from = share,
    values_fill = 0
  ) %>%
  rename(
    cheap_share     = Cheap,
    mid_share       = Mid,
    expensive_share = Expensive
  )

final_canton_data <- final_canton_data %>%
  left_join(canton_segments, by = "canton")

data_final_swiss <- data_final_swiss %>%
  mutate(
    product_group = case_when(
      str_detect(category, "smartphone")                    ~ "Smartphone",
      str_detect(category, "gaming")                        ~ "Gaming",
      str_detect(category, "sportuhr-smartwatch")           ~ "Smartwatch",
      str_detect(category, "einbaukuehlschrank|foehn")      ~ "Home/Appliance",
      TRUE                                                  ~ "Other"
    )
  )

group_shares <- data_final_swiss %>%
  group_by(canton, product_group) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(canton) %>%
  mutate(share = n / sum(n)) %>%
  ungroup() %>%
  select(canton, product_group, share) %>%
  pivot_wider(
    names_from  = product_group,
    values_from = share,
    values_fill = 0
  ) %>%
  rename(
    smartphone_share = Smartphone,
    gaming_share     = Gaming,
    smartwatch_share = Smartwatch,
    home_share       = `Home/Appliance`
  )

final_canton_data <- final_canton_data %>%
  left_join(group_shares, by = "canton")

canton_map_data <- canton_sf %>%
  left_join(final_canton_data, by = "canton")
```

### 7.1.2 Average Digitec Transaction Price by Canton

```{r}
#| label: interactive-canton-meanprice
#| warning: false
#| message: false

library(dplyr)
library(ggplot2)
library(plotly)
library(scales)

canton_map_data <- canton_map_data %>%
  mutate(
    tooltip_mean = paste0(
      "Canton: ", canton,
      "<br>Mean order value: ", round(avg_order_value, 1), " CHF",
      "<br>Total sales: ", round(total_sales, 0), " CHF",
      "<br>Transactions: ", num_transactions,
      "<br>Sales per capita: ", round(sales_per_capita, 2), " CHF"
    )
  )

p_mean <- ggplot(canton_map_data) +
  geom_sf(
    aes(
      fill = avg_order_value,
      text = tooltip_mean           
    ),
    color = "grey20",
    linewidth = 0.2
  ) +
  scale_fill_viridis_c(
    option = "C",
    name = "Mean order value [CHF]"
  ) +
  labs(
    title = "Average Digitec Transaction Price by Canton",
  ) +
  theme_minimal()

ggplotly(p_mean, tooltip = "text")


```

*“Where do customers spend the most per order?”*

This interactive map shows large regional differences in average order value. Cantons like AI, SZ, ZG and OW show significantly higher spending levels per transaction compared to the national average.

**Marketing implication for Digitec**

Understanding where customers tend to place expensive orders helps identify regions with greater purchasing power and a stronger need for premium electronics. These insights help Digitec to focus premium advertising and exclusive product promotions in the regions where customers are most willing to invest in high-end devices.

Target premium campaigns and up-sell messaging to high-value cantons to maximize revenue per customer.

### 7.1.3  Digitec Transactions per Canton

```{r}

p_canton_trans <- ggplot(canton_map_data) +
  geom_sf(
    aes(
      fill = num_transactions,
      text = paste0(
        "Canton: ", canton,
        "<br>Total transactions: ", num_transactions,
        "<br>Total sales: ", round(total_sales, 0), " CHF",
        "<br>Sales per capita: ", round(sales_per_capita, 2), " CHF",
        "<br>Mean order value: ", round(avg_order_value, 1), " CHF"
      )
    ),
    colour = "grey30",
    linewidth = 0.05
  ) +
  scale_fill_viridis_c(
    option = "C",
    na.value = "white",
    name = "Transactions"
  ) +
  labs(
    title = "Digitec transactions per canton"
  ) +
  theme_minimal()

library(plotly)
ggplotly(p_canton_trans, tooltip = "text")

```

*“Where is Digitec most active?”*

This map shows the number of purchases made across Switzerland. Cantons with large populations and strong digital adoption, such as Zurich or Bern, naturally generate more transactions, while others show lower activity.

**Marketing implication for Digitec**

Transaction volume highlights where Digitec has its strongest market presence and where further growth is possible. In cantons with low activity, Digitec may need to advertise more or explain delivery options better so that more people start shopping there.

It is best to use the very active cantons such as Zurich or Bern for big product launches, because many people will see them. Cantons with fewer purchases should receive more basic marketing to grow awareness and trust.

### 7.1.4  Share of smartphone-related purchases by canton

Smartphones are one of Digitec’s strongest product categories: high demand and frequent upgrades. They represent a “universal tech product” bought by almost every demographic group. In short, almost everyone buys a phone, and people replace them often. This makes smartphones a good way to understand general buying behaviour.\

Understanding regional differences in smartphone buying helps Digitec target the right audiences with mobile phone deals, subscription bundles or upgrade

```{r}
#| label: interactive-canton-smartphone-share
#| warning: false
#| message: false

canton_map_data <- canton_map_data %>%
  mutate(
    tooltip_smart = paste0(
      "Canton: ", canton,
      "<br>Smartphone share: ",
      percent(smartphone_share, accuracy = 0.1),
      "<br>Male share: ", percent(share_male, accuracy = 0.1),
      "<br>Avg order value: ", round(avg_order_value, 1), " CHF"
    )
  )

p_smart <- ggplot(canton_map_data) +
  geom_sf(aes(fill = smartphone_share),
          colour = "grey20",
          linewidth = 0.2) +
  scale_fill_viridis_c(
    option = "C",
    labels = percent_format(accuracy = 1),
    name   = "Smartphone purchase share"
  ) +
  labs(
    title = "Where do people buy smartphones?",
  ) +
  theme_minimal()

ggplotly(p_smart, tooltip = "tooltip_smart")

```

*“Where do people buy smartphones most often?”*

This visualization shows the distribution of smartphone purchases across cantons, with some regions like Jura and Graubünden showing exceptionally high smartphone demand.

**Marketing implication for Digitec**

Smartphones are one of Digitec’s most important products, and people replace them regularly. Knowing which regions buy the most smartphones helps Digitec decide where to advertise new phone models, upgrades, and accessories.

The cantons like Jura and Graubünden with the highest smartphone shares are the best places to run phone-focused promotions for Digitec.

### 7.1.5 Gaming Purchase Share by Canton

Gaming is a very strong and growing market segment for Digitec. It includes consoles, graphics cards, monitors and gaming PCs which are all high-value items. Gaming purchases also strongly correlate with demographics (younger, tech-oriented and often male). This makes gaming an excellent category for targeted marketing because interest varies significantly across regions.

```{r}
#| label: interactive-canton-gaming
#| warning: false
#| message: false

canton_map_data <- canton_map_data %>%
  mutate(
    tooltip_gaming = paste0(
      "Canton: ", canton,
      "<br>Gaming share: ", percent(gaming_share, accuracy = 0.1),
      "<br>Male share: ", percent(share_male, accuracy = 0.1),
      "<br>Avg order value: ", round(avg_order_value, 1), " CHF"
    )
  )

p_gaming <- ggplot(canton_map_data) +
  geom_sf(aes(fill = gaming_share, text = tooltip_gaming),
          colour = "grey20",
          linewidth = 0.2) +
  scale_fill_viridis_c(
    option = "C",
    labels = percent_format(accuracy = 1),
    name   = "Gaming\npurchase share"
  ) +
  labs(
    title = "Where do people buy gaming-related products?",
  ) +
  theme_minimal()

ggplotly(p_gaming, tooltip = "text")

```

This map highlights regions where gaming-related products (such as consoles, GPUs, and monitors) are most frequently purchased.

**Marketing implication for Digitec**

Gaming customers often display high engagement and repeated purchasing behavior. Knowing where gaming demand is strongest helps Digitec deliver campaigns that resonate with these audiences, especially during product launches, major game releases, and seasonal sales.

Digitec should focus gaming events, influencer collaborations and bundle promotions in top gaming cantons like Neuchâtel, Glarus and Valais.

### 7.1.6 Gender Composition by Canton

```{r}
#| label: interactive-canton-maleshare
#| warning: false
#| message: false

canton_map_data <- canton_map_data %>%
  mutate(
    tooltip_gender = paste0(
      "Canton: ", canton,
      "<br>Male share: ", percent(share_male, accuracy = 0.1),
      "<br>Avg order value: ", round(avg_order_value, 1), " CHF",
      "<br>Gaming share: ", percent(gaming_share, accuracy = 0.1),
      "<br>Smartphone share: ", percent(smartphone_share, accuracy = 0.1)
    )
  )

p_gender <- ggplot(canton_map_data) +
  geom_sf(aes(fill = share_male, text = tooltip_gender),
          colour = "grey20",
          linewidth = 0.2) +
  scale_fill_viridis_c(
    option = "C",
    labels = percent_format(accuracy = 1),
    name   = "Male share"
  ) +
  labs(
    title = "Gender composition by canton",
    subtitle = "Share of male population and tech preferences (hover)"
  ) +
  theme_minimal()

ggplotly(p_gender, tooltip = "text")

```

*“Which regions are more male or female?”*

This map displays the share of male population per canton, with some regions leaning slightly male and others slightly female. When you interact with the map it's also possible to see the insights such as average order value, gaming share and smartphone share from the previous visualizations.

**Marketing implication for Digitec**

Even minor gender imbalances can shape regional purchasing behavior so it's very important for Digitec to take it into account. Male-dominant regions tend to show greater interest in categories such as gaming and wearables, while female-dominant regions lean toward home appliances and lifestyle electronics. These demographic differences provide a valuable insight for audience segmentation.

Digitec should adjust product recommendations and homepage content to reflect gender-driven preferences within each canton. Next visualization will show the difference in purchases made by different gender dominated cantons.

#### 7.1.7 What Do Male vs Female Regions Buy?

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)

gender_category_summary <- final_canton_data %>%
  select(canton, share_male) %>%
  mutate(
    gender_group = if_else(share_male >= median(share_male, na.rm = TRUE),
                           "male",
                           "female")
  ) %>%
  left_join(group_shares, by = "canton") %>%
  group_by(gender_group) %>%
  summarise(
    Smartphones   = mean(smartphone_share, na.rm = TRUE),
    Gaming        = mean(gaming_share, na.rm = TRUE),
    Smartwatches  = mean(smartwatch_share, na.rm = TRUE),
    `Home appliances` = mean(home_share, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = -gender_group,
    names_to = "category",
    values_to = "share"
  )

ggplot(gender_category_summary,
       aes(x = category, y = share, fill = gender_group)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
  labs(
    title = "What do male vs female regions typically buy?",
    x = "Product group",
    y = "Share of purchases",
    fill = "Region type"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 20, hjust = 1)
  )



```

This chart compares product preferences in male-dominant and female-dominant cantons. Although one might expect gaming products to be more popular in male regions, the data shows the opposite: female-dominant regions actually buy slightly more gaming items.

**Marketing implication for Digitec**

This is important because it shows that common assumptions about gender and technology shopping do not always match real behaviour. Digitec should rely on data rather than stereotypes when planning campaigns. Gaming promotions can be targeted in female-dominant areas just as effectively as in male ones and smartphone campaigns are likely to work well across both.

### 7.1.8 Single vs Married Regions and what Do They Buy.

```{r}
marital_profile <- final_canton_data %>%
  mutate(
    marital_group = if_else(share_single > share_married,
                            "Single-dominant",
                            "Married-dominant")
  ) %>%
  select(
    marital_group, smartphone_share, gaming_share,
    smartwatch_share, home_share
  ) %>%
  group_by(marital_group) %>%
  summarise(across(everything(), mean, na.rm = TRUE)) %>%
  pivot_longer(cols = -marital_group, names_to = "category", values_to = "share")

ggplot(marital_profile, aes(x = category, y = share, fill = marital_group)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = percent_format(0.1)) +
  labs(
    title = "What do single vs married regions buy?",
    x = "Product category",
    y = "Purchase share",
    fill = "Region type"
  ) +
  theme_minimal()
```

This visualization contrasts product preferences between regions dominated by singles and those dominated by married populations.

**Marketing implication for Digitec**

Household structure significantly influences purchasing motivations. Single regions buy slightly more gaming products, while married regions lean a bit more toward smartwatches. Overall, household structure does not strongly affect what people buy, but these small tendencies can help Digitec adjust its marketing. For example gaming works slightly better in single-dominant areas, and smartwatch or wellness-related promotions may perform better in married-dominant ones.



## 9. Monthly Purchase Activity

```{r}
#| label: monthly-seasonality

library(lubridate)

monthly_sales <- data_final_swiss %>%
  mutate(month = floor_date(dateTime, "month")) %>%
  group_by(month) %>%
  summarise(
    total_sales = sum(price, na.rm = TRUE),
    avg_price   = mean(price, na.rm = TRUE),
    transactions = n()
  )

ggplot(monthly_sales, aes(x = month, y = transactions)) +
  geom_line(color = "#0099C6", linewidth = 1.3) +
  geom_point(color = "#0099C6", size = 2) +
  labs(
    title = "Monthly purchase activity",,
    x = "Month",
    y = "Number of transactions"
  ) +
  theme_minimal()


```

*"When do people buy more?"*

This line plot shows transaction trends throughout the year. Even though the dataset ends in October, the pattern clearly indicates rising interest during late spring and summer months.

**Marketing implication for Digitec**

Seasonality strongly influences consumer electronics purchases. Understanding when demand naturally rises allows Digitec to plan promotions, stock allocation and marketing budgets more efficiently. Although holiday-season data in digitec set is missing, the observed patterns still provide useful insight into mid-year behavior.

Digitec can strengthen promotional campaigns before summer peaks and prepare inventory well in advance.
